---
title: "Check noun repetitions, continuations and question distributions"
author: "Miriam Schulz"
date: "2024-11-27"
output:
  html_document:
    number_sections: true
    toc: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# About

This script reads in a `csv` stimuli file for the Cloze Prediction & Production experiment as well as the fillers file
and checks for overlap of the nouns within and across both sets.


# Duplicate checks

## Check if there are duplicate nouns in the items (context and target nouns)

In a first step, the data is read in and the chosen items (indicated by a $1$ in the column `picked`) are subsetted:

```{r, message=FALSE}
library("tidyverse")
```

```{r cars}
# Read in the data
dat <- read.csv("stim_cand.csv", header=TRUE)

# Filter the data set for the chosen candidate items
dat <- filter(dat, candidates=="1")

# Print items that do not occur exactly twice
subset(as.data.frame(xtabs(~ item_id + candidates, dat)), Freq != 2)
```

Write a function that will check if a word appears more than once,
either in the context nouns, in the target nouns, or across context and target nouns:

```{r}
check_overlap <- function(data, col1, col2) {
  # Check duplicates within each column
  duplicates_col1_table <- sort(table(data[[col1]])[table(data[[col1]]) > 2], decreasing=TRUE) / 2
  duplicates_col1 <- names(duplicates_col1_table)
  print("Duplicates within the context nouns:")
  print(duplicates_col1_table)
  
  duplicates_col2_table <- sort(table(data[[col2]])[table(data[[col2]]) > 1], decreasing=TRUE)
  duplicates_col2 <- names(duplicates_col2_table)
  print("Duplicates in the target nouns:")
  print(duplicates_col2_table)
  
  # Check for words appearing in both columns
  #duplicates_all_table <- (table(data[[col1]]) / 2) + (table(data[[col2]]))
  table1 <- table(data[[col1]]) / 2
  table2 <- table(data[[col2]])
  all_names <- union(names(table1), names(table2))
  aligned_table1 <- table1[all_names]
  aligned_table2 <- table2[all_names]
  aligned_table1[is.na(aligned_table1)] <- 0
  aligned_table2[is.na(aligned_table2)] <- 0
  duplicates_all_table <- sort(aligned_table1 + aligned_table2, decreasing=TRUE)
  duplicates_all_table <- duplicates_all_table[!is.na(names(duplicates_all_table))]

  print("Duplicates across both context and target nouns:")
  print(duplicates_all_table)
  
  common_words <- intersect(data[[col1]], data[[col2]])
  
  # Return results as a list
  return(list(
    duplicates_in_col1 = duplicates_col1,
    duplicates_in_col2 = duplicates_col2,
    common_words = common_words
  ))
}
```

Print the deuplicate nouns:

```{r}
overlap <- check_overlap(dat, "W3", "response")
overlap
```


## Detect if a noun is still free to use in the fillers

Check if a noun has been used as context or target noun in the items, or if it is still free to use as a noun in the fillers:

```{r}
# Create a set of all nouns used as context or target noun in the items:
noun_set <- sort(unique(c(dat$W3, dat$response)))
# Check if a specific word is present in this set (TRUE/FALSE):
"mother" %in% noun_set
"work" %in% noun_set
```


## Check the fillers for overlap with the items

Once the fillers have been written, we should check again if there are any nouns in the final filler set that occur as context/target noun in the items.

Read in the file `fillers.csv` and write a function to check for overlap with the item noun set:

```{r}
# Read in the fillers
fillers <- read.csv("fillers.csv", header=TRUE)

# Merge the context sentence and target columns into a single vector
filler_sentences <- c(fillers$Sentence, fillers$Target)

# Split the sentences into individual words
filler_words <- unique(tolower(unlist(strsplit(filler_sentences, "\\s+"))))
filler_words <- filler_words[filler_words != ""]  # remove empty string
#head(filler_words)
```

```{r}
check_filler_overlap <- function(item_nouns, filler_words) {
  for (w in filler_words) {
    if (w %in% item_nouns) {
      print(w)
    }
  }
}

check_filler_overlap(noun_set, filler_words)
```


# Continuations

## Check the continuations (spillover regions)

How often is each continuation (spillover word 1 + spillover word 2) used?

```{r}
spill <- select(dat, Spill1, Spill2)

# Check both spillover words as combination
spill$Spillover <- paste(spill$Spill1, spill$Spill2, sep = " ")
spillover_table <- data.frame(sort(xtabs(~ Spillover, spill) / 2, decreasing = TRUE))
write.csv(spillover_table, "spillover_counts.csv", row.names=FALSE)

# Check first spillover word only
spill1_table <- data.frame(sort(xtabs(~ Spill1, spill) / 2, decreasing = TRUE))
write.csv(spill1_table, "spillover_counts_spill1.csv", row.names=FALSE)
```


# Comprehension questions

## Check the comprehension questions in the items

```{r}
dat$Sentence <- paste(dat$W1, dat$W2, dat$W3, dat$W4, dat$W5, 
                      dat$LastWord, dat$response,
                      dat$Spill1, dat$Spill2, dat$Continuation, dat$End,
                      sep = " ")
                      
questions <- select(dat, item_id, Sentence, HasQuestion, Question, Answer, QuestionAbout)

# Check that 1/3 of all items have a question:
length(unique(questions$item_id))  # N all items
xtabs(~ HasQuestion, questions) /2  # N items with question

# Subset
questions <- filter(questions, HasQuestion == "Yes")

# Check that there is a question about each sentence position equally often
xtabs(~ QuestionAbout, questions) / 2

# Check that each sentence position has a true vs. false answer equally often
xtabs(~ Answer + QuestionAbout, questions) / 2
```

Inspect the actual questions:

```{r}
questions <- arrange(questions, QuestionAbout, Answer, item_id)
questions$HasQuestion <- NULL
questions$Answer <- ifelse(questions$Answer == "D", "False", "True")
write.csv(questions, "comprehension_questions.csv", row.names=FALSE)
```


## Check the comprehension questions in the fillers

```{r}
fillers <- filter(fillers, !(FillerType %in% c("prac_comprehension", "prac_production")))
questions_fillers <- select(fillers, Filler_id, FillerType, UnexpectedNoun, Sentence, Target,
                            HasQuestion, Question, Answer, QuestionAbout)

# Check that 1/3 of all items have a question:
length(unique(questions_fillers$Filler_id))  # N all fillers
xtabs(~ HasQuestion, questions_fillers)  # N fillers with question

# Subset
questions_fillers <- filter(questions_fillers, HasQuestion == "Yes")

# Check that there is a question about each sentence position ~ equally often
xtabs(~ QuestionAbout, questions_fillers)

# Check that each sentence position has a true vs. false answer ~ equally often
xtabs(~ Answer + QuestionAbout, questions_fillers)

# Check that each filler type has equal amounts of true/false answers
xtabs(~ FillerType + Answer, questions_fillers)

# Check that there are some unexpected nouns with questions
# (but distribution won't be 50-50 bc there are fewer unexpected noun fillers)
xtabs(~ UnexpectedNoun, questions_fillers)
```

