---
title: "Power analysis"
author: "Miriam Schulz"
date: "2024-12-13"
output:
  html_document:
    number_sections: true
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width = 16)
```

```{r load_libraries, message = FALSE}
library(lme4)
library(lmerTest)
library(designr)
library(dplyr)
library(ggplot2)
rm(list = ls())
```

# About

Power analysis for the Cloze Prediction & Production experiment.


## Design

A 2 x 2 target manipulation design.

Dependent variable: RTs.

Predictor variables

- Expectancy (high vs. low)
- Task (Cloze production vs. reading comprehension only)

TODO: add more detailed description of how to use the script.


# A first simulation with exaggerated effects

## Creating the data set

```{r generate_first_dataset}
set.seed(42)

design <-
  fixed.factor("Expectancy", levels=c("High", "Low")) +
  fixed.factor("Task", levels=c("Production", "Comprehension")) +
  random.factor("Item", instances=20) +
  random.factor("Subj", instances=10) +
  random.factor(c("Subj", "Item"), groups=c("Expectancy", "Task"))
dat <- dplyr::arrange(design.codes(design), Subj, Item)
length(unique(dat$Item)) # number of items
length(unique(dat$Subj)) # number of subjects

# convert to factor
dat <- dat %>% mutate(across(all_of(colnames(dat)), ~ as.factor(.)))

# set contrasts
contrasts(dat$Expectancy) <- c(-0.5, 0.5)
contrasts(dat$Task) <- c(-0.5, 0.5)
```

## Simulating the response variable

Define the mean (intercept), effect size for each effect, and the random effects and residual variation, then use `simLMM()` to generate random data fulfilling these criteria as a response variable:

```{r generate_first_RTs}
# Specify fixed effects coefficients for:
# Intercept, Expectancy, Task, Exp:Task Interaction
fix <- c(250, 40, 20, 10)

# Specify random effects + res variance
sd_Subj <- c(20, 10, 10)
sd_Item <- c(10, 10, 10)
sd_Res  <- 50

dat$RTsim <- simLMM(form = ~ 1 + Expectancy * Task +
                      (1 + Expectancy + Task | Subj) + 
                      (1 + Expectancy + Task | Item),
                   data   = dat,                     # data
                   Fixef  = fix,                     # fixed effects 
                   VC_sd  = list(sd_Subj, sd_Item, sd_Res),   # random effects and noise (error) SD
                   CP = 0.3,                         # correlation parameters of the random effects 
                   empirical = TRUE)
```

Sanity check: Do the fixed effects in the data set correspond to the constraints defined above?
(If `empirical = TRUE` is used above, this should correspond _exactly_ to the constaints.)

```{r verify_group_means}
dat %>% group_by(Expectancy) %>% summarize(M=mean(RTsim))
dat %>% group_by(Task) %>% summarize(M=mean(RTsim))
```

Run an LMM to verify that this in effect yields the right estimates:

```{r first_model, eval=FALSE}
# Interaction model
m_interaction <- lmer(RTsim ~ 1 + Expectancy * Task +
                      (1 + Expectancy + Task || Subj) +
                      (1 + Expectancy + Task || Item),
          data = dat, control=lmerControl(calc.derivs=FALSE))

save(m_interaction, file="model_test.rda")
```

```{r}
load(file="model_test.rda")
summary(m_interaction)
```

# Power analysis


## Run power analysis

To simulate many data sets, write a function that generates a new data frame for a given number of items and subjects:

```{r dataframe_generation_function}
generate_dataframe <- function(n_subj, n_items) {
  
  # Generate Latin Square design
  design <-
    fixed.factor("Expectancy", levels=c("High", "Low")) +
    fixed.factor("Task", levels=c("Production", "Comprehension")) +
    random.factor("Item", instances=n_items/4) +
    random.factor("Subj", instances=n_subj/4) +
    random.factor(c("Subj", "Item"), groups=c("Expectancy", "Task"))
  dat <- dplyr::arrange(design.codes(design), Subj, Item)

  # Convert to factor
  dat <- dat %>% mutate(across(all_of(colnames(dat)), ~ as.factor(.)))

  # Set contrasts
  contrasts(dat$Expectancy) <- c(-0.5, 0.5)
  contrasts(dat$Task) <- c(-0.5, 0.5)
  
  return(dat)
}
```

Define a function for the power analysis:

```{r power_analysis_function}
power_analysis <- function(dat,
                           mean_rt,
                           expectancy_fixed, task_fixed, interaction_fixed,
                           n_subj, n_items,
                           verbose=FALSE) {
  
  for (s in 1:length(n_subj)) {
    
    cat("SIMULATION NO. ", s, "/", length(n_subj), sep=" ", end="\n")
    
    for (i in 1:length(n_items)) {
      
      dat <- generate_dataframe(n_subj[s], n_items[i])
      nsj <- length(unique(dat$Subj))
      nit <- length(unique(dat$Item))
      
      if (verbose == TRUE) {
        cat("   Unique subjects:", length(unique(dat$Subj)), sep=" ", end="\n")
        cat("   Unique items:   ", length(unique(dat$Item)), sep=" ", end="\n")
      }
      
      for (e in 1:length(expectancy_fixed)) {


        # Extract the fixed effects for both factors
        exp <- expectancy_fixed[e]
        task <- task_fixed[e]
        interact <- interaction_fixed[e]
        
        if (verbose == TRUE) {
          cat("      Expectancy:", exp,
              "- Task:", task,
              "- Interaction:", interact,
              sep=" ", end="\n")
        }

        fix <- c(mean_rt, exp, task, interact)

        # For each iteration, run several simulations for more stable results
        for (j in 1:10) {

          # Simulate the response variable
          dat$RTsim <- simLMM(
            form = ~ 1 + Expectancy * Task +
              (1 + Expectancy + Task | Subj) +
              (1 + Expectancy + Task | Item),
            data   = dat,
            Fixef  = fix,
            VC_sd  = list(sd_Subj, sd_Item, sd_Res),
            CP     = 0.3,
            empirical = FALSE,
            verbose = FALSE
          )
          
          #ww <- ""  # init
          
          # Run an LMM
          suppressMessages(
            suppressWarnings(LMM <- withCallingHandlers({
              lmer(
                RTsim ~ 1 + Expectancy * Task +
                  # (1 + Expectancy + Task || Subj) +
                  # (1 + Expectancy + Task || Item),
                  (1 | Subj) +
                  (1 | Item),
                data = dat,
                REML = FALSE,
                control = lmerControl(calc.derivs = FALSE)
              )
            }#,
            #warning = function(w) { ww <<- w$message }
          )))
          
          # Append the LMM estimates to the COF data frame
          LMMcof <- coef(summary(LMM))
          COF <- rbind(
            COF, 
            c(nsj, nit, exp, task, interact,
              LMMcof["Expectancy1", ], LMMcof["Task1", ],
              LMMcof["Expectancy1:Task1", ],
              isSingular(LMM))
          )
          #warn[s] <- ww
        }
      }
    }
  }
  #COF$warning <- warn
  return(COF)
}
```

Set the fixed and random effects, and initialize:

```{r set_effects_init}
# fix <- c(250, 8, 5)   # Fixed effects: Intercept, Expectancy, Task

# Define a mean RT (intercept)
mean_rt <- 250  # intercept

# Define fixed effects for Expectancy and Task (vectors of same length)
# and their interaction.
# Note that these vectorsmust be the SAME length, but one can vary 
# which fixed effect to test by keeping the values in the vectors for the other
# effects constant, e.g., task_fixed <- c(5, 5, 5).
expectancy_fixed <- c(10, 10, 10)  # simulation values for expectancy
task_fixed <- c(5, 5, 5)  # simulation values for task
interaction_fixed <- c(0, 5, 10)  # simulation values for exp:task interaction

# Define random effects and variance
sd_Subj <- c(30, 10, 10)
sd_Item <- c(30, 10, 10)
sd_Res  <- 50

# Define vectors for the number of subjects and items to simulate
# n_subj <- seq(8, 96, 8)
# n_items <- seq(80, 96, 8)
n_subj <- seq(8, 96, 8)
n_items <- seq(80, 96, 16)

COF <- data.frame()
#warn <- c()
```

Run power analysis:

```{r power_analysis}
# set.seed(666)
set.seed(42)
COF <- power_analysis(dat,
                      mean_rt,
                      expectancy_fixed, task_fixed, interaction_fixed,
                      n_subj, n_items,
                      verbose = TRUE)
#warn <- power_output[2]
```

Format power analysis data frame:

```{r format_power_df}
# Results for LMMs
names(COF) <- c("nsj", "nit", "exp", "task", "interact",
                "Expectancy", "SE1", "df1", "t1", "p1",
                "Task", "SE2", "df2", "t2", "p2", 
                "Interaction", "SE3", "df3", "t3", "p3", 
                "singular")
#COF$warning <- warn
#COF$noWarning <- warn==""
COF$sign1   <- as.numeric(COF$p1 < .05) # determine significant results; one-tailed
COF$sign2   <- as.numeric(COF$p2 < .05) # Significant results for Task
COF$sign3   <- as.numeric(COF$p3 < .05) # Significant results for Task

COF$nsjF   <- gtools::quantcut(COF$nsj, q=seq(0,1,length=10))
COF$nsjFL  <- plyr::ddply(COF,"nsjF",transform,nsjFL=mean(nsj))$nsjFL

#save(COF, file="power.rda")
```


## Plot power results

### Power by number of subjects

Plot results by number of subjects (averaging over the simulations by N items):

```{r plot_subjects}
# Extract maximal y axis range to make the two plots comparable side by side
y_range <- range(c(COF$sign1, COF$sign2), na.rm = TRUE)

# Plot only the maximal fixed effects here
COF_large_effects <- filter(COF, exp == max(exp))
COF_large_effects <- filter(COF, task == max(task))
COF_large_effects <- filter(COF, interact == max(interact))

power_plot <- function(dat, sign, y_range, plt_title) {
  p <- ggplot(data = dat) +
    geom_smooth(aes(x = nsj, y = eval(parse(text=sign))), method="loess") +
    geom_point(   stat = "summary", aes(x = nsjFL, y = eval(parse(text=sign))),
                  fun.data = mean_se) +
    geom_errorbar(stat = "summary", aes(x = nsjFL, y = eval(parse(text=sign))),
                  fun.data = mean_se) +
    geom_line(    stat = "summary", aes(x = nsjFL, y = eval(parse(text=sign))),
                  fun.data = mean_se) +
    geom_hline(yintercept = 0.8, color = "red", linetype = "dashed") +
    geom_hline(yintercept = 1, color = "darkgreen", linetype = "solid") +
    ggtitle(plt_title) + 
    theme_minimal()
  p <- p + coord_cartesian(ylim = y_range)
  return(p)
}

p1 <- power_plot(COF_large_effects, "sign1", y_range, "Power for Expectancy")
p2 <- power_plot(COF_large_effects, "sign2", y_range, "Power for Task")
p3 <- power_plot(COF_large_effects, "sign3", y_range, "Power for the Interaction")

# # Visualize results for Expectancy
# p1 <- ggplot(data = COF_large_effects) +
#   geom_smooth(aes(x = nsj, y = sign1), method="loess") +
#   geom_point(   stat = "summary", aes(x = nsjFL, y = sign1), fun.data = mean_se) +
#   geom_errorbar(stat = "summary", aes(x = nsjFL, y = sign1), fun.data = mean_se) +
#   geom_line(    stat = "summary", aes(x = nsjFL, y = sign1), fun.data = mean_se) +
#   geom_hline(yintercept = 0.8, color = "red", linetype = "dashed") +
#   geom_hline(yintercept = 1, color = "darkgreen", linetype = "solid") +
#   ggtitle("Power for Expectancy") +
#   theme_minimal()
# 
# # Visualize results for Task
# p2 <- ggplot(data = COF_large_effects) +
#   geom_smooth(aes(x = nsj, y = sign2), method="loess") +
#   geom_point(   stat = "summary", aes(x = nsjFL, y = sign2), fun.data = mean_se) +
#   geom_errorbar(stat = "summary", aes(x = nsjFL, y = sign2), fun.data = mean_se) +
#   geom_line(    stat = "summary", aes(x = nsjFL, y = sign2), fun.data = mean_se) +
#   geom_hline(yintercept = 0.8, color = "red", linetype = "dashed") +
#   geom_hline(yintercept = 1, color = "darkgreen", linetype = "solid") +
#   ggtitle("Power for Task") +
#   theme_minimal()
# 
# # Visualize results for the Expectancy:Task interaction
# p3 <- ggplot(data = COF_large_effects) +
#   geom_smooth(aes(x = nsj, y = sign3), method="loess") +
#   geom_point(   stat = "summary", aes(x = nsjFL, y = sign3), fun.data = mean_se) +
#   geom_errorbar(stat = "summary", aes(x = nsjFL, y = sign3), fun.data = mean_se) +
#   geom_line(    stat = "summary", aes(x = nsjFL, y = sign3), fun.data = mean_se) +
#   geom_hline(yintercept = 0.8, color = "red", linetype = "dashed") +
#   geom_hline(yintercept = 1, color = "darkgreen", linetype = "solid") +
#   ggtitle("Power for the Interaction") +
#   theme_minimal()
# 
# p1 <- p1 + coord_cartesian(ylim = y_range)
# p2 <- p2 + coord_cartesian(ylim = y_range)
# p3 <- p3 + coord_cartesian(ylim = y_range)

suppressMessages({
  gridExtra::grid.arrange(p1, p2, p3, nrow = 1)
})
```


### Power by number of items

Plot by number of subjects + items:

```{r plot_items}
p1 <- ggplot(data=COF_large_effects) +
  geom_smooth(aes(x=nsj, y=sign1, colour=factor(nit), fill=factor(nit)),
              method="loess", se=FALSE)+
  geom_point(   stat="summary", aes(x=nsjFL, y=sign1, colour=factor(nit)))+
  geom_errorbar(stat="summary", aes(x=nsjFL, y=sign1, colour=factor(nit)))+
  geom_line(    stat="summary", aes(x=nsjFL, y=sign1, colour=factor(nit)))+
  geom_hline(yintercept = 0.8, color = "red", linetype = "dashed") +
  geom_hline(yintercept = 1, color = "darkgreen", linetype = "solid") +
  ggtitle("Power for Expectancy") + 
  theme_minimal()
  
p2 <- ggplot(data=COF_large_effects) +
  geom_smooth(aes(x=nsj, y=sign2, colour=factor(nit), fill=factor(nit)),
              method="loess", se=FALSE)+
  geom_point(   stat="summary", aes(x=nsjFL, y=sign2, colour=factor(nit)))+
  geom_errorbar(stat="summary", aes(x=nsjFL, y=sign2, colour=factor(nit)))+
  geom_line(    stat="summary", aes(x=nsjFL, y=sign2, colour=factor(nit)))+
  geom_hline(yintercept = 0.8, color = "red", linetype = "dashed") +
  geom_hline(yintercept = 1, color = "darkgreen", linetype = "solid") +
  ggtitle("Power for Task") + 
  theme_minimal()
  
p3 <- ggplot(data=COF_large_effects) +
  geom_smooth(aes(x=nsj, y=sign3, colour=factor(nit), fill=factor(nit)),
              method="loess", se=FALSE)+
  geom_point(   stat="summary", aes(x=nsjFL, y=sign3, colour=factor(nit)))+
  geom_errorbar(stat="summary", aes(x=nsjFL, y=sign3, colour=factor(nit)))+
  geom_line(    stat="summary", aes(x=nsjFL, y=sign3, colour=factor(nit)))+
  geom_hline(yintercept = 0.8, color = "red", linetype = "dashed") +
  geom_hline(yintercept = 1, color = "darkgreen", linetype = "solid") +
  ggtitle("Power for the Interaction") + 
  theme_minimal()

p1 <- p1 + coord_cartesian(ylim = y_range)
p2 <- p2 + coord_cartesian(ylim = y_range)
p3 <- p3 + coord_cartesian(ylim = y_range)

suppressMessages({
  gridExtra::grid.arrange(p1, p2, p3, nrow = 1)
})
```


### Power by effect size

Plot by number of subjects + effect size (only the max number of items):

```{r plot_effect_sizes}
COF_max_items <- filter(COF, nit == max(nit))

p1 <- ggplot(data=COF_max_items) +
  geom_smooth(aes(x=nsj, y=sign1, color=factor(exp), fill=factor(exp)),
              method="loess", se=TRUE)+
  geom_point(   stat="summary", aes(x=nsjFL, y=sign1, colour=factor(exp)))+
  geom_errorbar(stat="summary", aes(x=nsjFL, y=sign1, colour=factor(exp)))+
  geom_line(    stat="summary", aes(x=nsjFL, y=sign1, colour=factor(exp)))+
  geom_hline(yintercept = 0.8, color = "red", linetype = "dashed") +
  geom_hline(yintercept = 1, color = "darkgreen", linetype = "solid") +
  ggtitle("Power for Expectancy") + 
  theme_minimal()

p2 <- ggplot(data=COF_max_items) +
  geom_smooth(aes(x=nsj, y=sign2, color=factor(task), fill=factor(task)),
              method="loess", se=TRUE)+
  geom_point(   stat="summary", aes(x=nsjFL, y=sign2, colour=factor(task)))+
  geom_errorbar(stat="summary", aes(x=nsjFL, y=sign2, colour=factor(task)))+
  geom_line(    stat="summary", aes(x=nsjFL, y=sign2, colour=factor(task)))+
  geom_hline(yintercept = 0.8, color = "red", linetype = "dashed") +
  geom_hline(yintercept = 1, color = "darkgreen", linetype = "solid") +
  ggtitle("Power for Task") + 
  theme_minimal()

p3 <- ggplot(data=COF_max_items) +
  geom_smooth(aes(x=nsj, y=sign3, color=factor(interact), fill=factor(interact)),
              method="loess", se=TRUE)+
  geom_point(   stat="summary", aes(x=nsjFL, y=sign3, colour=factor(interact)))+
  geom_errorbar(stat="summary", aes(x=nsjFL, y=sign3, colour=factor(interact)))+
  geom_line(    stat="summary", aes(x=nsjFL, y=sign3, colour=factor(interact)))+
  geom_hline(yintercept = 0.8, color = "red", linetype = "dashed") +
  geom_hline(yintercept = 1, color = "darkgreen", linetype = "solid") +
  ggtitle("Power for the Interaction") + 
  theme_minimal()

p1 <- p1 + coord_cartesian(ylim = y_range)
p2 <- p2 + coord_cartesian(ylim = y_range)
p3 <- p3 + coord_cartesian(ylim = y_range)

suppressMessages({
  gridExtra::grid.arrange(p1, p2, p3, nrow = 1)
})
```



## N subjects needed to reach 80% power

Determine the number of subjects needed for each effect to reach a power of 80%:

```{r check_power_80%}
for (e in 1:length(expectancy_fixed)) {
  
  exp_current <- expectancy_fixed[e]
  task_current <- task_fixed[e]
  interact_current <- interaction_fixed[e]
  
  COF_current <- filter(COF, exp==exp_current)
  COF_current <- filter(COF_current, task==task_current)
  COF_current <- filter(COF_current, interact==interact_current)
  
  # Expectancy:
  m0 <- loess(sign1 ~ nsj, data=COF_current)
  COF$pred <- predict(m0)
  idx <- COF$pred>0.8
  min_exp <- min(COF$nsj[idx])
  
  cat("Expectancy effect size:", exp_current,
      " -> min required subjects:",
      min_exp, sep=" ", end="\n")

  # Task:
  m0 <- loess(sign2 ~ nsj, data=COF_current)
  COF$pred <- predict(m0)
  idx <- COF$pred>0.8
  min_task <- min(COF$nsj[idx])
  
  cat("Task effect size:", task_current,
      "        -> min required subjects:",
      min_task, sep=" ", end="\n")
  
  # Interaction:
  m0 <- loess(sign3 ~ nsj, data=COF_current)
  COF$pred <- predict(m0)
  idx <- COF$pred>0.8
  min_interact <- ifelse(length(COF$nsj[idx]) > 0, min(COF$nsj[idx]), FALSE)
  
  cat("Interaction effect size:", interact_current,
      "-> min required subjects:",
      min_interact, sep=" ", end="\n\n")
}

```

