---
title: "Analysis"
author: "Miriam Schulz"
date: "2025-01-08"
output:
  html_document:
    number_sections: true
    toc: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# About

This script contains the analysis for the Cloze Prediction & Production experiment.


# Preliminaries

Libraries & co:

```{r preliminaries, message = FALSE}
rm(list = ls())
library(tidyverse)
library(ggplot2)
library(RColorBrewer)
library(gridExtra)
library(Rmisc)
library(lme4)
library(lmerTest)
source("functions.R")
```


## Set flags 

Some script parameters (flags) that will determine which sections of the script will / will not be run:

```{r flags}
# Use only a subset of strong items with the minimal specified Cloze difference:
strong_item_subset = FALSE
cloze_difference <- 0.5

# Limit how many/which plots to generate:
generate_main_plots = TRUE
plot_inidividual_subjects = FALSE
plot_inidividual_items = FALSE

# Quit before running the models:
run_models = TRUE
```

## Load data

```{r read-data}
# Read preprocessed data
# dat <- read.csv("results_preprocessed.csv", header=TRUE)
load("results_preprocessed_targetregions.RData")
load("results_preprocessed_allregions.RData")
```


# Strong item subset analysis (optional)

Set `use_subset = TRUE` to look only at the "strongest" items in terms of Cloze probability in the entire following script.

```{r subset-strong-items}
if (strong_item_subset == TRUE) {
  
  dat_strong <- dat %>% 
    dplyr::group_by(Item, ExpCondition) %>%
    dplyr::summarise(mean_cloze = mean(ClozeProb), .groups = "drop") %>% # 
    pivot_wider(names_from = ExpCondition, values_from = mean_cloze) %>% # wide format
    mutate(difference = abs(High - Low)) %>% # Calculate the absolute difference
    filter(difference >= cloze_difference) %>% # Filter items with a difference >= 0.5
    select(Item)
  
  dat_strong <- dat %>% 
    filter(Item %in% dat_strong$Item) %>% 
    select(Item, ExpCondition, ClozeProb, TargetWord) %>% 
    unique() %>% 
    arrange(Item, ExpCondition)
  
  cat("FILTERING DATA SET FOR STRONG ITEMS ONLY\n")
  
  cat("Subsetting data frame for", length(unique(dat_strong$Item)),
      "unique items with a cloze probability difference of at least",
      cloze_difference, "between conditions.")
  
  dat <- dat %>% 
    filter(Item %in% dat_strong$Item)
}
```


# Participant subset analysis (optional)

TODO: add code from other script here



# RT Plots

## RTs main plot

```{r rt-line-plot, fig.height=5, fig.width=8}
# Generate a summary data frame including the SE:
dat_plot <- summarySE(dat,
                      measurevar="RT",
                      groupvars=c("Region", "ExpCondition", "Task"),
                      na.rm=T)

ggplot(data = dat_plot,
       aes(x = Region,
           y = RT,
           # color = interaction(Task, ExpCondition),
           color = ExpCondition,
           linetype = Task,
           group = ExpCondition)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2, show.legend = FALSE) +
  geom_errorbar(aes(ymin = RT - se, ymax = RT + se), 
                width = 0.2, 
                linewidth = 0.5) +
  facet_wrap(~Task) +
  xlab("Region") +
  ylab("Reading Time (ms)") +
  ggtitle("Reading Times by Task") +
  labs(color = "Expectancy") +
  scale_color_manual(values = c("navy", "steelblue1")) +
  # scale_color_manual(
  #   name = "Task and Condition", # Legend title
  #   values = c("Comprehension.High" = "navy",
  #              "Comprehension.Low" = "steelblue1",
  #              "Production.High" = "seagreen4",
  #              "Production.Low" = "#8de0a4"
  #              ),
  #   labels = c("HIGH - Comprehension", "LOW - Comprehension",
  #              "HIGH - Production", "LOW - Production"),
  #   breaks = c("Comprehension.High", "Comprehension.Low",
  #              "Production.High", "Production.Low"),
  #   guide = guide_legend(override.aes = list(linetype = c(1, 1, 2, 2)))
  # ) +
  # scale_linetype_manual(values = c("Comprehension" = "solid",
  #                                  "Production" = "dashed"),
  #                       guide = FALSE) +
  theme_minimal() +
  theme(
    axis.title.x = element_text(size = 14, face = "bold"),
    axis.title.y = element_text(size = 14, face = "bold"),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 12),
    legend.key.width = unit(1, "cm"),
    strip.text = element_text(size = 16, face = "bold"),
    plot.background = element_rect(fill = "white", color = NA)
  )

ggsave("./plots/RTs.png", plot = last_plot(),
       width=10, height=7, dpi=320)
```


## RTs by experiment half + task

```{r rt-line-plot-taskorder, eval = generate_main_plots, fig.height=5, fig.width=8}
# Generate a summary data frame including the SE:
dat_plot <- summarySE(dat,
                      measurevar="RT",
                      groupvars=c("Region", "ExpCondition",
                                  "Task", "Half"),
                      na.rm=T)

ggplot(data = dat_plot,
       aes(x = Region,
           y = RT,
           color = ExpCondition,
           linetype = Task,
           group = interaction(ExpCondition, Task))) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  geom_errorbar(aes(ymin = RT - se, ymax = RT + se), 
                width = 0.2, 
                linewidth = 0.5) +
  facet_wrap(~Half) +
  labs(color = "Expectancy") +
  xlab("Region") +
  ylab("Reading Time (ms)") +
  ggtitle("Reading Times by Region and Task Order") +
  theme_minimal() +
  theme(
    axis.title.x = element_text(size = 14, face = "bold"),
    axis.title.y = element_text(size = 14, face = "bold"),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 12),
    strip.text = element_text(size = 16, face = "bold"),
    plot.background = element_rect(fill = "white", color = NA)
  ) +
  scale_color_manual(values = c("navy", "steelblue1"))

ggsave("./plots/RTs_task_order.png", plot = last_plot(),
       width=10, height=7, dpi=320)
```


## RTs over time by subject

Plot the RT over time (items and fillers; averaging across the 4 extracted regions).

```{r rt-time-plot, eval = generate_main_plots, fig.height=7, fig.width=12}
dat_plot <- summarySE(dat_all,
                      measurevar="RT",
                      groupvars=c("Block", "TrialCounterGlobal",
                                  "UniqueID", "Task", "TaskOrder"),
                      na.rm=T)

ggplot(data = dat_plot,
       aes(x = TrialCounterGlobal,
           y = RT,
           color = UniqueID,
           linetype = Task,
           group = interaction(UniqueID, Block))) +
  geom_smooth(method = "lm", formula = y ~ x, se = FALSE, linewidth = 1.2) +
  # geom_line(linewidth = 1) +
  # geom_point(size = 2) +
  labs(color = "Subject", linetype = "Task") +
  xlab("Trial number") +
  ylab("Reading Time (ms)") +
  geom_vline(xintercept = 90, linetype = "solid", color = "black",
             linewidth = 0.8) +
  annotate("text", x = 87, y = 500, #y = max(dat_plot$RT-100),
           label = "Start second half",
           color = "black", angle = 90, hjust = 0.5) +
  ggtitle("Reading Times by Block") +
  facet_wrap(~TaskOrder) +
  theme_minimal() +
  theme(
    axis.title.x = element_text(size = 14, face = "bold"),
    axis.title.y = element_text(size = 14, face = "bold"),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 12),
    strip.text = element_text(size = 14, face = "bold"),
    plot.background = element_rect(fill = "white", color = NA)
  )

ggsave("./plots/RTs_over_time.png", plot = last_plot(),
       width=14, height=6, dpi=320)
```


## RTs by wordlength

```{r rt-targetlength-plot, eval = generate_main_plots}
# Generate a summary data frame including the SE:
dat_plot <- summarySE(filter(dat, !(Region %in% c("other"))),
                      measurevar="RT",
                      groupvars=c("TargetLength", "Region"),
                      na.rm=T)


# Determine minimal + maximal value for y-axis:
min_y <- min(c(dat_plot$RT - dat_plot$se))
max_y <- max(c(dat_plot$RT + dat_plot$se))

# Plot:
ggplot(data=dat_plot,
       aes(TargetLength, RT)) +
  geom_smooth(method = "lm", formula = y ~ x, se = TRUE, linewidth = 1.2, color = "steelblue") +
  geom_point(color = "steelblue", size = 3) +
  geom_errorbar(aes(ymin = RT-se, ymax = RT+se),
              width=0.1, linewidth=0.3) +
  xlab("Target Word Length in N characters") +
  ylab("Reading Time (ms)") +
  ggtitle("RT by Target Word Length") +
  ylim(min_y, max_y) +
  facet_wrap(~ Region) +
  theme_minimal() + 
  theme(
      axis.title.x = element_text(size = 14, face = 'bold'),
      axis.title.y = element_text(size = 14, face = 'bold'),
      axis.text.x = element_text(size = 12, face = 'bold'),
      axis.text.y = element_text(size = 12),
      plot.title = element_text(size = 25, face = 'bold', hjust = 0.5),
      legend.position = "none",
      strip.text = element_text(size = 18, face = "bold"),
      plot.background = element_rect(fill = "white", color = NA)
  )

ggsave("./plots/RTs_by_target_word_length.png", plot = last_plot(),
       width=12, height=6, dpi=320)
```


## RTs by target word frequency

```{r rt-targetfreq-plot, eval = generate_main_plots}
# Generate a summary data frame including the SE:
dat_plot <- dat %>% 
  filter(!(Region %in% c("other"))) %>% 
  mutate(TargetFreq = round(TargetFreq, 1))

dat_plot <- summarySE(dat_plot,
                      measurevar="RT",
                      groupvars=c("TargetFreq", "Region"),
                      na.rm=T)

# Determine minimal + maximal value for y-axis:
min_y <- min(c(dat_plot$RT - dat_plot$se))
max_y <- max(c(dat_plot$RT + dat_plot$se))

# Plot:
ggplot(data=dat_plot,
       aes(TargetFreq, RT)) +
  geom_smooth(method = "lm", formula = y ~ x, se = TRUE,
              linewidth = 1.2, color = "steelblue") +
  geom_point(color = "steelblue", size = 3) +
  geom_errorbar(aes(ymin = RT-se, ymax = RT+se),
              width=0.2, linewidth=0.3) +
              # position = position_dodge(0.8)) +
  xlab("Target Word Frequency (log10 SUBTLEX, rounded to 1 decimal)") +
  ylab("Reading Time (ms)") +
  ggtitle("RT by Target Word Frequency") +
  ylim(min_y, max_y) +
  facet_wrap(~ Region) +
  theme_minimal() + 
  theme(
      axis.title.x = element_text(size = 14, face = 'bold'),
      axis.title.y = element_text(size = 14, face = 'bold'),
      axis.text.x = element_text(size = 12, face = 'bold'),
      axis.text.y = element_text(size = 12),
      plot.title = element_text(size = 25, face = 'bold', hjust = 0.5),
      legend.position = "none",
      strip.text = element_text(size = 18, face = "bold"),
      plot.background = element_rect(fill = "white", color = NA)
  )

ggsave("./plots/RTs_by_target_word_freq.png", plot = last_plot(),
       width=12, height=6, dpi=320)
```


## Plot all subjects (in one plot)

```{r rt-subj-plot, eval = generate_main_plots, fig.height=15, fig.width=15}
dat_plot <- summarySE(dat,
                      measurevar="RT",
                      groupvars=c("Region", "ExpCondition",
                                  "Task", "UniqueID"),
                      na.rm=T)

ggplot(data = dat_plot,
       aes(x = Region,
           y = RT,
           color = ExpCondition,
           linetype = Task,
           group = interaction(ExpCondition, Task))) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  geom_errorbar(aes(ymin = RT - se, ymax = RT + se), 
                width = 0.2, 
                linewidth = 0.5) +
  facet_wrap(~UniqueID) +
  labs(color = "Expectancy") +
  xlab("Region") +
  ylab("Reading Time (ms)") +
  ggtitle("Reading Times by Subject") +
  theme_minimal() +
  theme(
    axis.title.x = element_text(size = 14, face = "bold"),
    axis.title.y = element_text(size = 14, face = "bold"),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 12),
    strip.text = element_text(size = 14, face = "bold"),
    plot.background = element_rect(fill = "white", color = NA)
  ) +
  scale_color_manual(values = c("navy", "steelblue1"))

ggsave("./plots/RTs_subjects.png", plot = last_plot(),
       width=25, height=13, dpi=320)
```


## Export individual subject plots

Generate and save one plot for each item, with the words annotated above the x-axis:

```{r subject-plots-individual}
if (plot_inidividual_subjects == TRUE) {
  dat_plot <- summarySE(dat,
                        measurevar="RT",
                        groupvars=c("Region", "ExpCondition",
                                    "Task", "UniqueID", "Subject"),
                        na.rm=T)
  
  for (i in unique(dat_plot$UniqueID)) {

    dat_plot_subj <- dat_plot %>% 
      filter(UniqueID == i)
    subj_no_char <- as.character(unique(dat_plot_subj$UniqueID))
    
    p <- ggplot(dat_plot_subj,
                aes(x = Region,
                y = RT,
                color = ExpCondition,
                linetype = Task,
                group = interaction(ExpCondition, Task))) +
      
      
      geom_line(linewidth = 1) +
      geom_point(size = 2) +
      geom_errorbar(aes(ymin = RT - se, ymax = RT + se), 
                    width = 0.2, 
                    linewidth = 0.5) +
      labs(color = "Expectancy",
           title = paste0("Reading Times: Subject ", subj_no_char)) +
      xlab("Region") +
      ylab("Reading Time (ms)") +
      theme_minimal() +
      theme(
        axis.title.x = element_text(size = 14, face = "bold"),
        axis.title.y = element_text(size = 14, face = "bold"),
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
        legend.title = element_text(size = 14),
        legend.text = element_text(size = 12),
        strip.text = element_text(size = 14, face = "bold"),
        plot.background = element_rect(fill = "white", color = NA)
      ) +
      scale_color_manual(values = c("navy", "steelblue1"))
    
    plot_title <- paste0("./plots/subject_plots/subject_",
                         as.character(i),
                         ".png")
    ggsave(plot_title, plot = p, width=12, height=5, dpi=320)
  }
}
```


## Plot all items (in one plot)

```{r rt-item-plot, eval = generate_main_plots, fig.height=25, fig.width=25}
# Generate a summary data frame including the SE:
dat_plot <- summarySE(dat,
                      measurevar="RT",
                      groupvars=c("Region", "ExpCondition",
                                  "Task", "ExpItemNum"),
                      na.rm=T)

ggplot(data = dat_plot,
       aes(x = Region,
           y = RT,
           color = ExpCondition,
           linetype = Task,
           group = interaction(ExpCondition, Task))) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  geom_errorbar(aes(ymin = RT - se, ymax = RT + se), 
                width = 0.2, 
                linewidth = 0.5) +
  # facet_wrap(~ExpItemNum+Task) +
  facet_wrap(~ExpItemNum) +
  labs(color = "Expectancy") +
  xlab("Region") +
  ylab("Reading Time (ms)") +
  ggtitle("Reading Times by Item") +
  theme_minimal() +
  theme(
    axis.title.x = element_text(size = 14, face = "bold"),
    axis.title.y = element_text(size = 14, face = "bold"),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 12),
    strip.text = element_text(size = 14, face = "bold"),
    plot.background = element_rect(fill = "white", color = NA)
  ) +
  scale_color_manual(values = c("navy", "steelblue1"))

ggsave("./plots/RTs_items.png", plot = last_plot(),
       width=30, height=20, dpi=320)
```


## Export individual item plots

Generate and save one plot for each item, with the words annotated above the x-axis:

```{r item-plots-individual}
if (plot_inidividual_items == TRUE) {
  dat_plot <- summarySE(dat_all,
                        measurevar="RT",
                        groupvars=c("Region", "ExpCondition",
                                    "Task", "Item",
                                    "WordPosition", "Word"),
                        na.rm=T)
  dat_plot$WordPosition <- as.numeric(dat_plot$WordPosition)
  summary(dat_plot$WordPosition)
  
  # for (i in c(1:80, 1001:1080)) {
  for (i in unique(as.numeric(as.character(dat_plot$Item)))) {
    dat_plot_item <- dat_plot %>% 
      filter(Item == i) %>% 
      group_by(WordPosition) %>%
      mutate(Word = if_else(
        WordPosition == 6,
        paste(unique(Word[WordPosition == 6]), collapse = "/"),
        Word)) %>%
      ungroup()
    item_no_char <- as.character(unique(dat_plot_item$Item))
    
    p <- ggplot(dat_plot_item,
                aes(x = WordPosition,
                y = RT,
                color = ExpCondition,
                linetype = Task,
                group = interaction(ExpCondition, Task))) +
      geom_line(linewidth = 1) +
      geom_point(size = 2) +
      geom_errorbar(aes(ymin = RT - se, ymax = RT + se), 
                    width = 0.2, 
                    linewidth = 0.5) +
      {if (as.numeric(item_no_char) < 1000)
           annotate("rect", xmin = 5.5, xmax = 8.5, ymin = -Inf, ymax = Inf, 
               fill = "gray", alpha = 0.3)} +
      labs(color = "Expectancy",
           title = paste0("Reading Times: Item ", item_no_char)) +
      scale_x_continuous(
        name = "Word Position",
        breaks = dat_plot_item$WordPosition,
        sec.axis = sec_axis(~ ., name = "Word",
                            breaks = dat_plot_item$WordPosition,
                            labels = dat_plot_item$Word)
      ) +
      ylab("Reading Time (ms)") +
      theme_minimal() +
      theme(
        axis.title.x = element_text(size = 14, face = "bold"),
        axis.title.y = element_text(size = 14, face = "bold"),
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
        legend.title = element_text(size = 14),
        legend.text = element_text(size = 12),
        strip.text = element_text(size = 14, face = "bold"),
        plot.background = element_rect(fill = "white", color = NA)
      ) +
      scale_color_manual(values = c("navy", "steelblue1"))
    
    plot_title <- paste0("./plots/item_plots/item_", as.character(i), ".png")
    ggsave(plot_title, plot = p, width=12, height=5, dpi=320)
  }
}
```

```{r exit-before-models, include = FALSE}
if (run_models == FALSE) {
  knitr::knit_exit()
}
```



# Models

Standardize numeric predictors, set contrasts and level order for categorical predictors, etc.

```{r prepare-predictors}
dat$Expectancy <- droplevels(factor(dat$ExpCondition,
                                    levels = c("High", "Low")))
contrasts(dat$Expectancy) <- c(-0.5, 0.5)
colnames(attr(dat$Expectancy, "contrasts")) <- c("Low")  # rename factor level

dat$Task <- factor(dat$Task,
                   levels = c("Comprehension", "Production"))
contrasts(dat$Task) <- c(-0.5, 0.5)
colnames(attr(dat$Task, "contrasts")) <- c("Production")  # rename factor level

dat$Half <- factor(dat$Half, levels = c("First Half", "Second Half"))
contrasts(dat$Half) <- c(-0.5, 0.5)
colnames(attr(dat$Half, "contrasts")) <- c("Second")  # rename factor level

dat$TaskOrder <- factor(dat$TaskOrder, levels = c("comp_prod", "prod_comp"))
contrasts(dat$TaskOrder) <- c(-0.5, 0.5)
colnames(attr(dat$TaskOrder, "contrasts")) <- c("Prod<Comp")  # rename factor level

dat$BlocksReversed <- factor(dat$BlocksReversed, levels = c("False", "True"))
contrasts(dat$BlocksReversed) <- c(-0.5, 0.5)
colnames(attr(dat$BlocksReversed, "contrasts")) <- c("True")  # rename factor level

dat$logRT <- log(dat$RT)
dat$logPreCritRT <- scale(log(dat$PreCriticalRT))
dat$Cloze_c <- scale(dat$ClozeProb)
dat$Length_c <- scale(dat$TargetLength)
dat$Freq_c <- scale(dat$TargetFreq)

dat$TrialCounterTask_c <- scale(dat$TrialCounterTask)
dat$TrialCounterGlobal_c <- scale(dat$TrialCounterGlobal)

#TODO: residualize for word length + frequency instead?
#TODO: add precritical RT as predictor?
```


## Target word

Lmer with cloze as predictor:

```{r lmer-target-cloze, eval = FALSE}
m_target_cloze <- lmer(logRT ~ Cloze_c * Task + Length_c + Freq_c +
                   TrialCounterTask_c + logPreCritRT +
                   (1 + Cloze_c + Task + TrialCounterTask_c | Subject) +
                   (1 + Cloze_c + Task | Item),
                 data = filter(dat, Region == "target"))
summary(m_target_cloze)
```

Lmer with binary expectancy as predictor:

```{r lmer-target-binary-expectancy, eval = FALSE}
m_target <- lmer(logRT ~ Expectancy * Task + Length_c + Freq_c +
                   (1 + Expectancy + Task | Subject) +
                   (1 + Expectancy + Task | Item),
                 data = filter(dat, Region == "target"))
summary(m_target)
```


## Spillover region 1

Lmer with cloze as predictor:

```{r lmer-spillover1-cloze, eval = FALSE}
m_spill1_cloze <- lmer(logRT ~ Cloze_c * Task + Length_c + Freq_c +
                   TrialCounterTask_c +
                   logPreCritRT +
                   (1 + Cloze_c + Task + TrialCounterTask_c | Subject) +
                   (1 + Cloze_c + Task | Item),
                 data = filter(dat, Region == "spill1"))
summary(m_spill1_cloze)
```

```{r lmer-spillover1-cloze-half, eval = FALSE}
# A main effect of half
lmer(logRT ~ Cloze_c + Task + Length_c + Freq_c + 
       Half +
       (1 | Subject) +
       (1 | Item),
     data = filter(dat, Region == "spill1")) %>% summary()

# But: this goes away when adding the random slope for Half, suggesting
# that this effect hinges on some participants:
lmer(logRT ~ Cloze_c + Task + Length_c + Freq_c + 
       Half +
       (1 + Half | Subject) +
       (1 | Item),
     data = filter(dat, Region == "spill1")) %>% summary()

# Reversed block order should have no effect
lmer(logRT ~ Cloze_c + Task + Length_c + Freq_c + 
       BlocksReversed +
       (1 | Subject) +
       (1 | Item),
     data = filter(dat, Region == "spill1")) %>% summary()
```


Lmer with binary expectancy as predictor:

```{r  lmer-spillover1-binary-expectancy, eval = FALSE}
# Spillover 1
m_spill1 <- lmer(logRT ~ Expectancy * Task + Length_c + Freq_c +
                   (1 + Expectancy + Task | Subject) +
                   (1 + Expectancy + Task | Item),
                 data = filter(dat, Region == "spill1"))
summary(m_spill1)
```


## Spillover region 2

Lmer with cloze as predictor:

```{r lmer-spillover2-cloze, eval = FALSE}
m_spill2_cloze <- lmer(logRT ~ Cloze_c * Task + Length_c + Freq_c +
                 TrialCounterTask_c + logPreCritRT +
                   (1 + Cloze_c + Task + TrialCounterTask_c | Subject) +
                   (1 + Cloze_c + Task | Item),
                 data = filter(dat, Region == "spill2"))
summary(m_spill2_cloze)
```

Lmer with binary expectancy as predictor:

```{r lmer-spillover2-binary-expectancy, eval = FALSE}
m_spill2 <- lmer(logRT ~ Expectancy * Task + Length_c + Freq_c +
                   (1 + Expectancy + Task | Subject) +
                   (1 + Expectancy + Task | Item),
                 data = filter(dat, Region == "spill2"))
summary(m_spill2)
```


## Plot model coefficients

```{r coef-plot}
get_coefficients <- function(m) {
  coef_data <- coef(summary(m)) %>%
  as.data.frame() %>%
  rownames_to_column(var = "term") %>%
  filter(term != "(Intercept)") %>% 
  dplyr::rename(
    estimate = Estimate,
    std.error = `Std. Error`,
    statistic = `t value`,
    p.value = `Pr(>|t|)`
  )
  return(coef_data)
}

dat_coef <- NULL
# for (region in c("pre-critical", "target", "spill1", "spill2")) {
for (region in c("target", "spill1", "spill2")) {
  dat_region <- dat %>% filter(Region == region)
  
  # m <- lmer(logRT ~ Expectancy * Task + Length_c + Freq_c +
  #             (1 | Subject) +
  #             (1 | Item),
  #           data = dat_region)
  
  m <- lmer(logRT ~ Cloze_c + Task + Cloze_c:Task +
              Half + Half:Task +
              Length_c + Freq_c +
                   TrialCounterTask_c +
                   logPreCritRT +
                   (1 + Cloze_c + Task + Half | Subject) +
                   # (1 + Cloze_c + Task + Half + Length_c + Freq_c | Subject) +
                   (1 | Item),
            data = dat_region)
  
  print("---------------------------------------------------------------------")
  print("MODEL SUMMARY FOR REGION:")
  print(region)
  print(summary(m))
            
  coef_region <- get_coefficients(m)
  coef_region$Region <- region
  
  if (!exists("dat_coef")) {
    dat_coef <- coef_region
    } else {
    dat_coef <- rbind(dat_coef, coef_region)
  }
}

dat_coef$Region <- factor(dat_coef$Region, levels =
                            c("pre-critical","target",
                              "spill1", "spill2"))

ggplot(filter(dat_coef, Region != "pre-critical"),
       aes(x = Region, y = estimate,
           color = term)) +
  geom_errorbar(aes(ymin = estimate - std.error,
                    ymax = estimate + std.error,
                    color = term),
                width=0.1, linewidth=0.3) +
  geom_point(aes(shape = p.value < 0.05), size = 3) +
  geom_hline(yintercept = 0, color = "black") +
  scale_shape_manual(values = c(13, 19)) +
  scale_color_brewer(palette = "Set1") +
  labs(title = "Model Coefficients",
       x = "Region",
       y = "Coefficient Estimate",
       color = "Predictor") +
  theme_minimal()
```

Try model with merged spillover regions:

```{r lmer-merged-spillover}
m_spillover_merged <- lmer(logRT ~ Cloze_c + Task + Cloze_c:Task + Region +
            Half + Half:Task +
            Length_c + Freq_c +
                 TrialCounterTask_c +
                 logPreCritRT +
                 (1 + Cloze_c + Task + Half + Region | Subject) +
                 # (1 + Cloze_c + Task + Half + Length_c + Freq_c | Subject) +
                 (1 | Item),
          data = filter(dat, Region %in% c("spill1", "spill2")))
  
print(summary(m_spillover_merged))
```

