---
title: "Analysis"
author: "Miriam Schulz"
date: "2025-01-08"
output:
  html_document:
    number_sections: true
    toc: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# About

This script contains the analysis for the Cloze Prediction & Production experiment.


# Preliminaries

Libraries & co:

```{r preliminaries, message = FALSE}
rm(list = ls())
library(tidyverse)
library(ggplot2)
library(RColorBrewer)
library(gridExtra)
library(Rmisc)
library(lme4)
library(lmerTest)
source("functions.R")
```


## Set flags 

Some script parameters (flags) that will determine which sections of the script will / will not be run:

```{r flags}
# Use only a subset of strong items with the minimal specified Cloze difference:
strong_item_subset <- FALSE
cloze_difference <- 0.5

# Limit how many/which plots to generate:
generate_main_plots <- TRUE
plot_inidividual_subjects <- TRUE
plot_inidividual_items<- TRUE
plot_inidividual_items_surprisal <- TRUE

# Run models:
run_models <- FALSE

# Run surprisal analyses:
run_surprisal_models<- FALSE
```

Set the plot colors:

```{r define-plot-colors}
plot_colors_conditions <- c("navy", "magenta3")
```


## Load data

```{r read-data}
# Read preprocessed data
# dat <- read.csv("results_preprocessed.csv", header=TRUE)
load("results_preprocessed_targetregions.RData")
load("results_preprocessed_allregions.RData")
```

# Subset analyses (optional)

## Strong item subset

Set `strong_item_subset = TRUE` in the flags at the top of this script to look only at the "strongest" items in terms of Cloze probability in the entire following script.

```{r subset-strong-items}
if (strong_item_subset == TRUE) {
  
  dat_strong <- dat %>% 
    dplyr::group_by(Item, ExpCondition) %>%
    dplyr::summarise(mean_cloze = mean(ClozeProb), .groups = "drop") %>% # 
    pivot_wider(names_from = ExpCondition, values_from = mean_cloze) %>% # wide format
    mutate(difference = abs(High - Low)) %>% # Calculate the absolute difference
    filter(difference >= cloze_difference) %>% # Filter items with a difference >= 0.5
    select(Item)
  
  dat_strong <- dat %>% 
    filter(Item %in% dat_strong$Item) %>% 
    select(Item, ExpCondition, ClozeProb, TargetWord) %>% 
    unique() %>% 
    arrange(Item, ExpCondition)
  
  cat("FILTERING DATA SET FOR STRONG ITEMS ONLY\n")
  
  cat("Subsetting data frame for", length(unique(dat_strong$Item)),
      "unique items with a cloze probability difference of at least",
      cloze_difference, "between conditions.")
  
  dat <- dat %>% 
    filter(Item %in% dat_strong$Item)
}
```


## Participant subset

TODO: add code from other script here



# RT Plots

## RTs main plot

```{r rt-line-plot, fig.height=5, fig.width=8}
# Generate a summary data frame including the SE:
dat_plot <- summarySE(dat,
                      measurevar="RT",
                      groupvars=c("Region", "ExpCondition", "Task"),
                      na.rm=T)

ggplot(data = dat_plot,
       aes(x = Region,
           y = RT,
           # color = interaction(Task, ExpCondition),
           color = ExpCondition,
           linetype = Task,
           group = ExpCondition)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2, show.legend = FALSE) +
  geom_errorbar(aes(ymin = RT - se, ymax = RT + se), 
                width = 0.2, 
                linewidth = 0.5) +
  facet_wrap(~Task) +
  xlab("Region") +
  ylab("Reading Time (ms)") +
  ggtitle("Reading Times by Task") +
  labs(color = "Expectancy") +
  # scale_color_manual(values = c("navy", "steelblue1")) +
  scale_color_manual(values = plot_colors_conditions) +
  # scale_color_manual(
  #   name = "Task and Condition", # Legend title
  #   values = c("Comprehension.High" = "navy",
  #              "Comprehension.Low" = "steelblue1",
  #              "Production.High" = "seagreen4",
  #              "Production.Low" = "#8de0a4"
  #              ),
  #   labels = c("HIGH - Comprehension", "LOW - Comprehension",
  #              "HIGH - Production", "LOW - Production"),
  #   breaks = c("Comprehension.High", "Comprehension.Low",
  #              "Production.High", "Production.Low"),
  #   guide = guide_legend(override.aes = list(linetype = c(1, 1, 2, 2)))
  # ) +
  # scale_linetype_manual(values = c("Comprehension" = "solid",
  #                                  "Production" = "dashed"),
  #                       guide = FALSE) +
  theme_minimal() +
  theme(
    axis.title.x = element_text(size = 14, face = "bold"),
    axis.title.y = element_text(size = 14, face = "bold"),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 12),
    legend.key.width = unit(1, "cm"),
    strip.text = element_text(size = 16, face = "bold"),
    plot.background = element_rect(fill = "white", color = NA)
  )

ggsave("./plots/RTs_twopanes.png", plot = last_plot(),
       width=9, height=6, dpi=320)
```

New plot: all in one pane:

```{r rt-line-plot-oneplot, fig.height=5, fig.width=8}
# Generate a summary data frame including the SE:
dat_plot <- summarySE(dat,
                      measurevar="RT",
                      groupvars=c("Region", "ExpCondition", "Task"),
                      na.rm=T)

ggplot(data = dat_plot,
       aes(x = Region,
           y = RT,
           # color = interaction(Task, ExpCondition),
           color = ExpCondition,
           linetype = Task,
           group = interaction(ExpCondition, Task))) +
  geom_line(linewidth = 1) +
  geom_point(size = 2, show.legend = FALSE) +
  geom_errorbar(aes(ymin = RT - se, ymax = RT + se), 
                width = 0.2, 
                linewidth = 0.5) +
  #facet_wrap(~Task) +
  xlab("Region") +
  ylab("Reading Time (ms)") +
  ggtitle("Reading Times by Task") +
  labs(color = "Expectancy") +
  # scale_color_manual(values = c("navy", "steelblue1")) +
  scale_color_manual(values = plot_colors_conditions) +
  theme_minimal() +
  theme(
    axis.title.x = element_text(size = 14, face = "bold"),
    axis.title.y = element_text(size = 14, face = "bold"),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 12),
    legend.key.width = unit(1, "cm"),
    strip.text = element_text(size = 16, face = "bold"),
    plot.background = element_rect(fill = "white", color = NA)
  )

ggsave("./plots/RTs.png", plot = last_plot(),
       width=6, height=4, dpi=320)
```

```{r rt-line-plot-oneplot-colorbytask, fig.height=5, fig.width=8}
# Generate a summary data frame including the SE:
dat_plot <- summarySE(dat,
                      measurevar="RT",
                      groupvars=c("Region", "ExpCondition", "Task"),
                      na.rm=T)

ggplot(data = dat_plot,
       aes(x = Region,
           y = RT,
           color = Task,
           linetype = ExpCondition,
           group = interaction(ExpCondition, Task))) +
  geom_line(linewidth = 1) +
  geom_point(size = 2, show.legend = FALSE) +
  geom_errorbar(aes(ymin = RT - se, ymax = RT + se), 
                width = 0.2, 
                linewidth = 0.5) +
  #facet_wrap(~Task) +
  xlab("Region") +
  ylab("Reading Time (ms)") +
  ggtitle("Reading Times by Task") +
  labs(color = "Task",
       linetype = "Expectancy") +
  # scale_color_manual(values = c("navy", "steelblue1")) +
  # scale_color_manual(values = c("navy", "forestgreen")) +
  # scale_color_manual(values = c("navy", "magenta3")) +
  scale_color_manual(values = plot_colors_conditions) +
  theme_minimal() +
  theme(
    axis.title.x = element_text(size = 14, face = "bold"),
    axis.title.y = element_text(size = 14, face = "bold"),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 12),
    legend.key.width = unit(1, "cm"),
    strip.text = element_text(size = 16, face = "bold"),
    plot.background = element_rect(fill = "white", color = NA)
  )

# ggsave("./plots/RTs_oneplot_color_by_task.png", plot = last_plot(),
#        width=6, height=4, dpi=320)
# 
# ggsave("./plots/RTs_oneplot_color_navy_forestgreen.png", plot = last_plot(),
#        width=6, height=4, dpi=320)

ggsave("./plots/RTs_oneplot_color_navy_magenta3.png", plot = last_plot(),
       width=6, height=5, dpi=320)
```



## RTs by experiment half + task

```{r rt-line-plot-taskorder-half, eval = generate_main_plots, fig.height=5, fig.width=8}
# Generate a summary data frame including the SE:
dat_plot <- summarySE(dat,
                      measurevar="RT",
                      groupvars=c("Region", "ExpCondition",
                                  "Task", "Half"),
                      na.rm=T)

ggplot(data = dat_plot,
       aes(x = Region,
           y = RT,
           color = ExpCondition,
           linetype = Task,
           group = interaction(ExpCondition, Task))) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  geom_errorbar(aes(ymin = RT - se, ymax = RT + se), 
                width = 0.2, 
                linewidth = 0.5) +
  facet_wrap(~Half) +
  labs(color = "Expectancy") +
  xlab("Region") +
  ylab("Reading Time (ms)") +
  ggtitle("Reading Times by Region and Experiment Half") +
  # scale_color_manual(values = c("navy", "steelblue1")) +
  scale_color_manual(values = plot_colors_conditions) +
  theme_minimal() +
  theme(
    axis.title.x = element_text(size = 14, face = "bold"),
    axis.title.y = element_text(size = 14, face = "bold"),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 12),
    strip.text = element_text(size = 16, face = "bold"),
    plot.background = element_rect(fill = "white", color = NA)
  )

ggsave("./plots/RTs_task_order.png", plot = last_plot(),
       width=10, height=7, dpi=320)
```


## RTs by block + task

```{r rt-line-plot-taskorder-block, eval = generate_main_plots, fig.height=5, fig.width=8}
# Generate a summary data frame including the SE:
dat_plot <- summarySE(dat,
                      measurevar="RT",
                      groupvars=c("Region", "ExpCondition",
                                  "Task", "Block"),
                      na.rm=T)

ggplot(data = dat_plot,
       aes(x = Region,
           y = RT,
           color = ExpCondition,
           linetype = Task,
           group = interaction(ExpCondition, Task))) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  geom_errorbar(aes(ymin = RT - se, ymax = RT + se), 
                width = 0.2, 
                linewidth = 0.5) +
  facet_wrap(~Block, nrow = 1) +
  labs(color = "Expectancy") +
  xlab("Region") +
  ylab("Reading Time (ms)") +
  ggtitle("Reading Times by Region and Block") +
  # scale_color_manual(values = c("navy", "steelblue1")) +
  scale_color_manual(values = plot_colors_conditions) +
  theme_minimal() +
  theme(
    axis.title.x = element_text(size = 14, face = "bold"),
    axis.title.y = element_text(size = 14, face = "bold"),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 12),
    strip.text = element_text(size = 16, face = "bold"),
    plot.background = element_rect(fill = "white", color = NA)
  )
ggsave("./plots/RTs_by_block.png", plot = last_plot(),
       width=15, height=7, dpi=320)
```



## RTs over time by subject

Plot the RT over time (items and fillers; averaging across the 4 extracted regions).

```{r rt-time-plot, eval = generate_main_plots, fig.height=7, fig.width=12}
dat_plot <- summarySE(dat_all,
                      measurevar="RT",
                      groupvars=c("Block", "TrialCounterGlobal",
                                  "UniqueID", "Task", "TaskOrder"),
                      na.rm=T)

ggplot(data = dat_plot,
       aes(x = TrialCounterGlobal,
           y = RT,
           color = UniqueID,
           linetype = Task,
           group = interaction(UniqueID, Block))) +
  geom_smooth(method = "lm", formula = y ~ x, se = FALSE, linewidth = 1.2) +
  # geom_line(linewidth = 1) +
  # geom_point(size = 2) +
  labs(color = "Subject", linetype = "Task") +
  xlab("Trial number") +
  ylab("Reading Time (ms)") +
  geom_vline(xintercept = 90, linetype = "solid", color = "black",
             linewidth = 0.8) +
  annotate("text", x = 87, y = 500, #y = max(dat_plot$RT-100),
           label = "Start second half",
           color = "black", angle = 90, hjust = 0.5) +
  ggtitle("Reading Times by Block") +
  facet_wrap(~TaskOrder) +
  theme_minimal() +
  theme(
    axis.title.x = element_text(size = 14, face = "bold"),
    axis.title.y = element_text(size = 14, face = "bold"),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 12),
    strip.text = element_text(size = 14, face = "bold"),
    plot.background = element_rect(fill = "white", color = NA)
  )

ggsave("./plots/RTs_over_time.png", plot = last_plot(),
       width=14, height=6, dpi=320)
```


## RTs by wordlength

```{r rt-targetlength-plot, eval = generate_main_plots}
# Generate a summary data frame including the SE:
dat_plot <- summarySE(filter(dat, !(Region %in% c("other"))),
                      measurevar="RT",
                      groupvars=c("TargetLength", "Region"),
                      na.rm=T)


# Determine minimal + maximal value for y-axis:
min_y <- min(c(dat_plot$RT - dat_plot$se))
max_y <- max(c(dat_plot$RT + dat_plot$se))

# Plot:
ggplot(data=dat_plot,
       aes(TargetLength, RT)) +
  geom_smooth(method = "lm", formula = y ~ x, se = TRUE, linewidth = 1.2, color = "steelblue") +
  geom_point(color = "steelblue", size = 3) +
  geom_errorbar(aes(ymin = RT-se, ymax = RT+se),
              width=0.1, linewidth=0.3) +
  xlab("Target Word Length in N characters") +
  ylab("Reading Time (ms)") +
  ggtitle("RT by Target Word Length") +
  ylim(min_y, max_y) +
  facet_wrap(~ Region) +
  theme_minimal() + 
  theme(
      axis.title.x = element_text(size = 14, face = 'bold'),
      axis.title.y = element_text(size = 14, face = 'bold'),
      axis.text.x = element_text(size = 12, face = 'bold'),
      axis.text.y = element_text(size = 12),
      plot.title = element_text(size = 25, face = 'bold', hjust = 0.5),
      legend.position = "none",
      strip.text = element_text(size = 18, face = "bold"),
      plot.background = element_rect(fill = "white", color = NA)
  )

ggsave("./plots/RTs_by_target_word_length.png", plot = last_plot(),
       width=12, height=6, dpi=320)
```


## RTs by target word frequency

```{r rt-targetfreq-plot, eval = generate_main_plots}
# Generate a summary data frame including the SE:
dat_plot <- dat %>% 
  filter(!(Region %in% c("other"))) %>% 
  mutate(TargetFreq = round(TargetFreq, 1))

dat_plot <- summarySE(dat_plot,
                      measurevar="RT",
                      groupvars=c("TargetFreq", "Region"),
                      na.rm=T)

# Determine minimal + maximal value for y-axis:
min_y <- min(c(dat_plot$RT - dat_plot$se))
max_y <- max(c(dat_plot$RT + dat_plot$se))

# Plot:
ggplot(data=dat_plot,
       aes(TargetFreq, RT)) +
  geom_smooth(method = "lm", formula = y ~ x, se = TRUE,
              linewidth = 1.2, color = "steelblue") +
  geom_point(color = "steelblue", size = 3) +
  geom_errorbar(aes(ymin = RT-se, ymax = RT+se),
              width=0.2, linewidth=0.3) +
              # position = position_dodge(0.8)) +
  xlab("Target Word Frequency (log10 SUBTLEX, rounded to 1 decimal)") +
  ylab("Reading Time (ms)") +
  ggtitle("RT by Target Word Frequency") +
  ylim(min_y, max_y) +
  facet_wrap(~ Region) +
  theme_minimal() + 
  theme(
      axis.title.x = element_text(size = 14, face = 'bold'),
      axis.title.y = element_text(size = 14, face = 'bold'),
      axis.text.x = element_text(size = 12, face = 'bold'),
      axis.text.y = element_text(size = 12),
      plot.title = element_text(size = 25, face = 'bold', hjust = 0.5),
      legend.position = "none",
      strip.text = element_text(size = 18, face = "bold"),
      plot.background = element_rect(fill = "white", color = NA)
  )

ggsave("./plots/RTs_by_target_word_freq.png", plot = last_plot(),
       width=12, height=6, dpi=320)
```


## Plot all subjects (in one plot)

```{r rt-subj-plot, eval = generate_main_plots, fig.height=15, fig.width=15}
dat_plot <- summarySE(dat,
                      measurevar="RT",
                      groupvars=c("Region", "ExpCondition",
                                  "Task", "UniqueID"),
                      na.rm=T)

ggplot(data = dat_plot,
       aes(x = Region,
           y = RT,
           color = ExpCondition,
           linetype = Task,
           group = interaction(ExpCondition, Task))) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  geom_errorbar(aes(ymin = RT - se, ymax = RT + se), 
                width = 0.2, 
                linewidth = 0.5) +
  facet_wrap(~UniqueID) +
  labs(color = "Expectancy") +
  xlab("Region") +
  ylab("Reading Time (ms)") +
  ggtitle("Reading Times by Subject") +
  # scale_color_manual(values = c("navy", "steelblue1")) +
  scale_color_manual(values = plot_colors_conditions) +
  theme_minimal() +
  theme(
    axis.title.x = element_text(size = 14, face = "bold"),
    axis.title.y = element_text(size = 14, face = "bold"),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 12),
    strip.text = element_text(size = 14, face = "bold"),
    plot.background = element_rect(fill = "white", color = NA)
  )

ggsave("./plots/RTs_subjects.png", plot = last_plot(),
       width=25, height=13, dpi=320)
```


## Export individual subject plots

Generate and save one plot for each item, with the words annotated above the x-axis:

```{r subject-plots-individual}
if (plot_inidividual_subjects == TRUE) {
  dat_plot <- summarySE(dat,
                        measurevar="RT",
                        groupvars=c("Region", "ExpCondition",
                                    "Task", "UniqueID", "Subject"),
                        na.rm=T)
  
  for (i in unique(dat_plot$UniqueID)) {

    dat_plot_subj <- dat_plot %>% 
      filter(UniqueID == i)
    subj_no_char <- as.character(unique(dat_plot_subj$UniqueID))
    
    p <- ggplot(dat_plot_subj,
                aes(x = Region,
                y = RT,
                color = ExpCondition,
                linetype = Task,
                group = interaction(ExpCondition, Task))) +
      
      
      geom_line(linewidth = 1) +
      geom_point(size = 2) +
      geom_errorbar(aes(ymin = RT - se, ymax = RT + se), 
                    width = 0.2, 
                    linewidth = 0.5) +
      labs(color = "Expectancy",
           title = paste0("Reading Times: Subject ", subj_no_char)) +
      xlab("Region") +
      ylab("Reading Time (ms)") +
      # scale_color_manual(values = c("navy", "steelblue1")) +
      scale_color_manual(values = plot_colors_conditions) +
      theme_minimal() +
      theme(
        axis.title.x = element_text(size = 14, face = "bold"),
        axis.title.y = element_text(size = 14, face = "bold"),
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
        legend.title = element_text(size = 14),
        legend.text = element_text(size = 12),
        strip.text = element_text(size = 14, face = "bold"),
        plot.background = element_rect(fill = "white", color = NA)
      )

    plot_title <- paste0("./plots/subject_plots/subject_",
                         as.character(i),
                         ".png")
    ggsave(plot_title, plot = p, width=12, height=5, dpi=320)
  }
}
```


## Plot all items (in one plot)

```{r rt-item-plot, eval = generate_main_plots, fig.height=25, fig.width=25}
# Generate a summary data frame including the SE:
dat_plot <- summarySE(dat,
                      measurevar="RT",
                      groupvars=c("Region", "ExpCondition",
                                  "Task", "ExpItemNum"),
                      na.rm=T)

ggplot(data = dat_plot,
       aes(x = Region,
           y = RT,
           color = ExpCondition,
           linetype = Task,
           group = interaction(ExpCondition, Task))) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  geom_errorbar(aes(ymin = RT - se, ymax = RT + se), 
                width = 0.2, 
                linewidth = 0.5) +
  # facet_wrap(~ExpItemNum+Task) +
  facet_wrap(~ExpItemNum) +
  labs(color = "Expectancy") +
  xlab("Region") +
  ylab("Reading Time (ms)") +
  ggtitle("Reading Times by Item") +
  # scale_color_manual(values = c("navy", "steelblue1")) +
  scale_color_manual(values = plot_colors_conditions) +
  theme_minimal() +
  theme(
    axis.title.x = element_text(size = 14, face = "bold"),
    axis.title.y = element_text(size = 14, face = "bold"),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 12),
    strip.text = element_text(size = 14, face = "bold"),
    plot.background = element_rect(fill = "white", color = NA)
  )

ggsave("./plots/RTs_items.png", plot = last_plot(),
       width=30, height=20, dpi=320)
```


## Export individual item plots

Generate and save one plot for each item, with the words annotated above the x-axis:

```{r item-plots-individual}
if (plot_inidividual_items == TRUE) {
  dat_plot <- summarySE(dat_all,
                        measurevar="RT",
                        groupvars=c("Region", "ExpCondition",
                                    "Task", "Item",
                                    "WordPosition", "Word"),
                        na.rm=T)
  dat_plot$WordPosition <- as.numeric(dat_plot$WordPosition)
  summary(dat_plot$WordPosition)
  
  # for (i in c(1:80, 1001:1080)) {
  for (i in unique(as.numeric(as.character(dat_plot$Item)))) {
    dat_plot_item <- dat_plot %>% 
      filter(Item == i) %>% 
      group_by(WordPosition) %>%
      mutate(Word = if_else(
        WordPosition == 6,
        paste(unique(Word[WordPosition == 6]), collapse = "/"),
        Word)) %>%
      ungroup()
    item_no_char <- as.character(unique(dat_plot_item$Item))
    
    p <- ggplot(dat_plot_item,
                aes(x = WordPosition,
                y = RT,
                color = ExpCondition,
                linetype = Task,
                group = interaction(ExpCondition, Task))) +
      geom_line(linewidth = 1) +
      geom_point(size = 2) +
      geom_errorbar(aes(ymin = RT - se, ymax = RT + se), 
                    width = 0.2, 
                    linewidth = 0.5) +
      {if (as.numeric(item_no_char) < 1000)
           annotate("rect", xmin = 5.5, xmax = 8.5, ymin = -Inf, ymax = Inf, 
               fill = "gray", alpha = 0.3)} +
      labs(color = "Expectancy",
           title = paste0("Reading Times: Item ", item_no_char)) +
      scale_x_continuous(
        name = "Word Position",
        breaks = dat_plot_item$WordPosition,
        sec.axis = sec_axis(~ ., name = "Word",
                            breaks = dat_plot_item$WordPosition,
                            labels = dat_plot_item$Word)
      ) +
      ylab("Reading Time (ms)") +
      # scale_color_manual(values = c("navy", "steelblue1")) +
      scale_color_manual(values = plot_colors_conditions) +
      theme_minimal() +
      theme(
        axis.title.x = element_text(size = 14, face = "bold"),
        axis.title.y = element_text(size = 14, face = "bold"),
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
        legend.title = element_text(size = 14),
        legend.text = element_text(size = 12),
        strip.text = element_text(size = 14, face = "bold"),
        plot.background = element_rect(fill = "white", color = NA)
      )

    plot_title <- paste0("./plots/item_plots/item_", as.character(i), ".png")
    ggsave(plot_title, plot = p, width=12, height=5, dpi=320)
  }
}
```

```{r exit-before-models, include = FALSE}
# if (run_models == FALSE) {
#   knitr::knit_exit()
# }
```



# Models

Standardize numeric predictors, set contrasts and level order for categorical predictors, etc.

```{r prepare-predictors}
dat$Expectancy <- droplevels(factor(dat$ExpCondition,
                                    levels = c("High", "Low")))
contrasts(dat$Expectancy) <- c(-0.5, 0.5)
colnames(attr(dat$Expectancy, "contrasts")) <- c("Low")  # rename factor level

dat$Task <- factor(dat$Task,
                   levels = c("Comprehension", "Production"))
contrasts(dat$Task) <- c(-0.5, 0.5)
colnames(attr(dat$Task, "contrasts")) <- c("Production")  # rename factor level

dat$Half <- factor(dat$Half, levels = c("First Half", "Second Half"))
contrasts(dat$Half) <- c(-0.5, 0.5)
colnames(attr(dat$Half, "contrasts")) <- c("Second")  # rename factor level

dat$Block <- as.factor(dat$Block)
contrasts(dat$Block) <- contr.sum(4)

dat$TaskOrder <- factor(dat$TaskOrder, levels = c("comp_prod", "prod_comp"))
contrasts(dat$TaskOrder) <- c(-0.5, 0.5)
colnames(attr(dat$TaskOrder, "contrasts")) <- c("Prod<Comp")  # rename factor level

dat$BlocksReversed <- factor(dat$BlocksReversed, levels = c("False", "True"))
contrasts(dat$BlocksReversed) <- c(-0.5, 0.5)
colnames(attr(dat$BlocksReversed, "contrasts")) <- c("True")  # rename factor level

dat$logRT <- log(dat$RT)
dat$logPreCritRT <- scale(log(dat$PreCriticalRT))
dat$Cloze_c <- scale(dat$ClozeProb)
dat$Length_c <- scale(dat$TargetLength)
dat$Freq_c <- scale(dat$TargetFreq)
dat <- dat %>% 
  mutate(WordLength = nchar(Word))
dat$TrialCounterTask_c <- scale(dat$TrialCounterTask)
dat$TrialCounterGlobal_c <- scale(dat$TrialCounterGlobal)

#TODO: residualize for word length + frequency instead?
#TODO: add precritical RT as predictor?
```


## Target word

Lmer with cloze as predictor:

```{r lmer-target-cloze, eval = FALSE}
m_target_cloze <- lmer(logRT ~ Cloze_c * Task + Length_c + Freq_c +
                   TrialCounterTask_c + logPreCritRT +
                   (1 + Cloze_c + Task + TrialCounterTask_c | Subject) +
                   (1 + Cloze_c + Task | Item),
                 data = filter(dat, Region == "target"))
summary(m_target_cloze)
```

Lmer with binary expectancy as predictor:

```{r lmer-target-binary-expectancy, eval = FALSE}
m_target <- lmer(logRT ~ Expectancy * Task + Length_c + Freq_c +
                   (1 + Expectancy + Task | Subject) +
                   (1 + Expectancy + Task | Item),
                 data = filter(dat, Region == "target"))
summary(m_target)
```


## Spillover region 1

Lmer with cloze as predictor:

```{r lmer-spillover1-cloze, eval = FALSE}
m_spill1_cloze <- lmer(logRT ~ Cloze_c * Task + Length_c + Freq_c +
                   TrialCounterTask_c +
                   logPreCritRT +
                   (1 + Cloze_c + Task + TrialCounterTask_c | Subject) +
                   (1 + Cloze_c + Task | Item),
                 data = filter(dat, Region == "spill1"))
summary(m_spill1_cloze)
```

```{r lmer-spillover1-cloze-half, eval = FALSE}
# A main effect of half
lmer(logRT ~ Cloze_c + Task + Length_c + Freq_c + 
       Half +
       (1 | Subject) +
       (1 | Item),
     data = filter(dat, Region == "spill1")) %>% summary()

# But: this goes away when adding the random slope for Half, suggesting
# that this effect hinges on some participants:
lmer(logRT ~ Cloze_c + Task + Length_c + Freq_c + 
       Half +
       (1 + Half | Subject) +
       (1 | Item),
     data = filter(dat, Region == "spill1")) %>% summary()

# Reversed block order should have no effect
lmer(logRT ~ Cloze_c + Task + Length_c + Freq_c + 
       BlocksReversed +
       (1 | Subject) +
       (1 | Item),
     data = filter(dat, Region == "spill1")) %>% summary()
```


Lmer with binary expectancy as predictor:

```{r lmer-spillover1-binary-expectancy, eval = FALSE}
# Spillover 1
m_spill1 <- lmer(logRT ~ Expectancy * Task + Length_c + Freq_c +
                   (1 + Expectancy + Task | Subject) +
                   (1 + Expectancy + Task | Item),
                 data = filter(dat, Region == "spill1"))
summary(m_spill1)
```


## Spillover region 2

Lmer with cloze as predictor:

```{r lmer-spillover2-cloze, eval = FALSE}
m_spill2_cloze <- lmer(logRT ~ Cloze_c * Task + Length_c + Freq_c +
                 TrialCounterTask_c + logPreCritRT +
                   (1 + Cloze_c + Task + TrialCounterTask_c | Subject) +
                   (1 + Cloze_c + Task | Item),
                 data = filter(dat, Region == "spill2"))
summary(m_spill2_cloze)
```

Lmer with binary expectancy as predictor:

```{r lmer-spillover2-binary-expectancy, eval = FALSE}
m_spill2 <- lmer(logRT ~ Expectancy * Task + Length_c + Freq_c +
                   (1 + Expectancy + Task | Subject) +
                   (1 + Expectancy + Task | Item),
                 data = filter(dat, Region == "spill2"))
summary(m_spill2)
```


## Plot model coefficients

```{r coef-plot, eval = run_models}
dat_coef <- NULL
# for (region in c("pre-critical", "target", "spill1", "spill2")) {
for (region in c("target", "spill1", "spill2")) {
  dat_region <- dat %>% filter(Region == region)
  
  # m <- lmer(logRT ~ Expectancy * Task + Length_c + Freq_c +
  #             (1 | Subject) +
  #             (1 | Item),
  #           data = dat_region)
  
  m <- lmer(logRT ~ Cloze_c + Task + Cloze_c:Task +
              Half + Half:Task +
              Length_c + Freq_c +
              #scale(WordLength) +
                   TrialCounterTask_c +
                   logPreCritRT +
                   (1 + Cloze_c + Task + Half | Subject) +
                   # (1 + Cloze_c + Task + Half + Length_c + Freq_c | Subject) +
                   (1 | Item),
            data = dat_region)
  
  print("---------------------------------------------------------------------")
  print("MODEL SUMMARY FOR REGION:")
  print(region)
  print(summary(m))
            
  coef_region <- get_coefficients(m)
  coef_region$Region <- region
  
  if (!exists("dat_coef")) {
    dat_coef <- coef_region
    } else {
    dat_coef <- rbind(dat_coef, coef_region)
  }
}

dat_coef$Region <- factor(dat_coef$Region, levels =
                            c("pre-critical","target",
                              "spill1", "spill2"))

ggplot(filter(dat_coef, Region != "pre-critical"),
       aes(x = Region, y = estimate,
           color = term)) +
  geom_errorbar(aes(ymin = estimate - std.error,
                    ymax = estimate + std.error,
                    color = term),
                width=0.1, linewidth=0.3) +
  geom_point(aes(shape = p.value < 0.05), size = 3) +
  geom_hline(yintercept = 0, color = "black") +
  scale_shape_manual(values = c(13, 19)) +
  scale_color_brewer(palette = "Set1") +
  labs(title = "Model Coefficients",
       x = "Region",
       y = "Coefficient Estimate",
       color = "Predictor") +
  theme_minimal() + 
    theme(
    axis.title.x = element_text(size = 14, face = "bold"),
    axis.title.y = element_text(size = 14, face = "bold"),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 12),
    strip.text = element_text(size = 14, face = "bold"),
    plot.background = element_rect(fill = "white", color = NA)
  )

ggsave("./plots/model_coefficients/model_coefficients.png", plot = last_plot(),
       width=10, height=7, dpi=320)
```

## Merged spillover regions model

Try model with merged spillover regions (simple su or mean).

NOTE THAT THESE MODELS WILL BE EQUIVALENT BESIDES A SLIGHT OFFSET IN THE INTERCEPT,
namely $−log(2)$, because $log(\frac{A}{B}) = log(A) - log(B)$:  
so for the mean, logRT will be $log(\frac{RTspill1 + RTspill2}{2}) = log(RTspill1 + RTspill2) - log(2)$,
while it is simply $log(RTspill1 + RTspill2)$ for the sum.  
Hence, the only difference is the constant offset $−log(2)$ for the mean, which only influences the intercept but does not impact the relationship between the predictors in the model.


```{r lmer-merged-spillover, eval = run_models}
dat_merged_spill <- dat %>%
  select(-Word, -WordPosition, -WordLength, -logRT) %>%  # Exclude the two columns
  pivot_wider(names_from = Region, values_from = RT, names_prefix = "RT")
dat_merged_spill$SumSpillRT <- ifelse(!is.na(dat_merged_spill$RTspill1) &
                                         !is.na(dat_merged_spill$RTspill2),
                                       dat_merged_spill$RTspill1 +
                                         dat_merged_spill$RTspill2,
                                       NA)
dat_merged_spill$MeanSpillRT <- dat_merged_spill$SumSpillRT / 2

dat_merged_spill$logSumSpillRT <- log(dat_merged_spill$SumSpillRT)
dat_merged_spill$logMeanSpillRT <- log(dat_merged_spill$MeanSpillRT)

m_spillover_sum <- lmer(logSumSpillRT ~ Cloze_c + Task + Cloze_c:Task +
                             Half + Half:Task +
                             Length_c + Freq_c + TrialCounterTask_c +
                             logPreCritRT +
                             (1 + Cloze_c + Task + Half | Subject) +
                             (1 | Item),
                           data = dat_merged_spill)
print(summary(m_spillover_sum))

m_spillover_mean <- lmer(logMeanSpillRT ~ Cloze_c + Task + Cloze_c:Task +
                             Half + Half:Task +
                             Length_c + Freq_c + TrialCounterTask_c +
                             logPreCritRT +
                             (1 + Cloze_c + Task + Half | Subject) +
                             (1 | Item),
                           data = dat_merged_spill)
print(summary(m_spillover_mean))
```


# Surprisal analyses

```{r read-surprisals}
surprisals_gpt2_large <- read.csv("stimuli_surprisal_gpt2-large.csv", header = TRUE)
surprisals_gpt_neo <- read.csv("stimuli_surprisal_gpt-neo.csv", header = TRUE)
surprisals_llama <- read.csv("stimuli_surprisal_llama.csv", header = TRUE)

surprisals_gpt2_large <- surprisals_gpt2_large %>% 
  select(c("ItemNum", "ExpCondition", "TargetWord",
           "TargetSurprisal", "Surprisals", "Probs")) %>% 
  dplyr::rename(ExpItemNum = ItemNum,
                TargetSurprisalGpt2Large = TargetSurprisal,
                SurprisalsGpt2Large = Surprisals,
                ProbsGpt2Large = Probs)

surprisals_gpt_neo <- surprisals_gpt_neo %>% 
  select(c("ItemNum", "ExpCondition", "TargetWord",
           "TargetSurprisal", "Surprisals", "Probs")) %>% 
  dplyr::rename(ExpItemNum = ItemNum,
                TargetSurprisalGptNeo = TargetSurprisal,
                SurprisalsGptNeo = Surprisals,
                ProbsGptNeo = Probs)

surprisals_llama <- surprisals_llama %>% 
  select(c("ItemNum", "ExpCondition", "TargetWord",
           "TargetSurprisal", "Surprisals", "Probs")) %>% 
  dplyr::rename(ExpItemNum = ItemNum,
                TargetSurprisalLlama = TargetSurprisal,
                SurprisalsLlama = Surprisals,
                ProbsLlama= Probs)

dat_surprisals <- merge(dat, surprisals_gpt2_large,
                        by = c("ExpItemNum", "ExpCondition", "TargetWord"))
dat_surprisals <- merge(dat_surprisals, surprisals_gpt_neo,
                        by = c("ExpItemNum", "ExpCondition", "TargetWord"))
dat_surprisals <- merge(dat_surprisals, surprisals_llama,
                        by = c("ExpItemNum", "ExpCondition", "TargetWord"))

dat_surprisals_all <- merge(dat_all, surprisals_gpt2_large,
                        by = c("ExpItemNum", "ExpCondition", "TargetWord"))
dat_surprisals_all <- merge(dat_surprisals_all, surprisals_gpt_neo,
                        by = c("ExpItemNum", "ExpCondition", "TargetWord"))#
dat_surprisals_all <- merge(dat_surprisals_all, surprisals_llama,
                        by = c("ExpItemNum", "ExpCondition", "TargetWord"))

dat_surprisals_all <- dat_surprisals_all %>% 
  mutate(WordLength = nchar(Word))
```

## Correlation between Surprisal and Cloze

```{r correlation-surprisal-cloze}
dat_surprisals_mini <- dat_surprisals %>%
  select("ExpItemNum", "ExpCondition", "TargetWord", "ClozeProb",
         "TargetSurprisalGpt2Large", "TargetSurprisalGptNeo",
         "TargetSurprisalLlama") %>% 
  unique()

# Pearson
cor(dat_surprisals_mini$ClozeProb,
    dat_surprisals_mini$TargetSurprisalGpt2Large)
cor(dat_surprisals_mini$ClozeProb,
    dat_surprisals_mini$TargetSurprisalGptNeo)
cor(dat_surprisals_mini$ClozeProb,
    dat_surprisals_mini$TargetSurprisalLlama)

# Spearman
cor(dat_surprisals_mini$ClozeProb,
    dat_surprisals_mini$TargetSurprisalGpt2Large,
    method = "spearman")
cor(dat_surprisals_mini$ClozeProb,
    dat_surprisals_mini$TargetSurprisalGptNeo,
    method = "spearman")
cor(dat_surprisals_mini$ClozeProb,
    dat_surprisals_mini$TargetSurprisalLlama,
    method = "spearman")

# Correlations among the surprisal values themselves (3 pairs)
cor(dat_surprisals_mini$TargetSurprisalGpt2Large,
    dat_surprisals_mini$TargetSurprisalGptNeo)
cor(dat_surprisals_mini$TargetSurprisalGpt2Large,
    dat_surprisals_mini$TargetSurprisalLlama)
cor(dat_surprisals_mini$TargetSurprisalGptNeo,
    dat_surprisals_mini$TargetSurprisalLlama)
# Makes sense: highest cor between the models closest in terms of N params.

ggplot(dat_surprisals_mini, aes(x = ClozeProb,
                                y = TargetSurprisalGpt2Large,
                                color = ExpCondition
                                )) +
  geom_point(size = 3) + 
  geom_text(aes(label = TargetWord), vjust = -0.5, size = 4) +
  geom_smooth(method = "lm", formula = y ~ x, color = "darkred", se = TRUE) +
  labs(
    title = "Correlation between Cloze Probability and Surprisal",
    subtitle = "gpt2-large, 812M parameters",
    x = "Target Cloze Probability",
    y = "Target Surprisal (gpt2-large)",
    color = "Condition"
  ) +
  # scale_color_manual(values = c("navy", "steelblue1")) +
  scale_color_manual(values = plot_colors_conditions) +
  theme_minimal() +
  theme(
    axis.title.x = element_text(size = 14, face = "bold"),
    axis.title.y = element_text(size = 14, face = "bold"),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 16, face = "italic", hjust = 0.5),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 12),
    strip.text = element_text(size = 14, face = "bold"),
    plot.background = element_rect(fill = "white", color = NA)
  )

ggsave("./plots/correlations/Surprisal_cloze_correlation_gpt2-large.png", plot = last_plot(),
       width=10, height=8, dpi=320)

ggplot(dat_surprisals_mini, aes(x = ClozeProb,
                                y = TargetSurprisalGptNeo,
                                color = ExpCondition
                                )) +
  geom_point(size = 3) + 
  geom_text(aes(label = TargetWord), vjust = -0.5, size = 4) +
  geom_smooth(method = "lm", formula = y ~ x, color = "darkred", se = TRUE) +
  labs(
    title = "Correlation between Cloze Probability and Surprisal",
    subtitle = "GPT-Neo, 2.72B parameters",
    x = "Target Cloze Probability",
    y = "Target Surprisal (GPT-Neo)",
    color = "Condition"
  ) +
  # scale_color_manual(values = c("navy", "steelblue1")) +
  scale_color_manual(values = plot_colors_conditions) +
  theme_minimal() +
  theme(
    axis.title.x = element_text(size = 14, face = "bold"),
    axis.title.y = element_text(size = 14, face = "bold"),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 16, face = "italic", hjust = 0.5),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 12),
    strip.text = element_text(size = 14, face = "bold"),
    plot.background = element_rect(fill = "white", color = NA)
  )

ggsave("./plots/correlations/Surprisal_cloze_correlation_gpt-neo.png", plot = last_plot(),
       width=10, height=8, dpi=320)

ggplot(dat_surprisals_mini, aes(x = ClozeProb,
                                y = TargetSurprisalLlama,
                                color = ExpCondition
                                )) +
  geom_point(size = 3) + 
  geom_text(aes(label = TargetWord), vjust = -0.5, size = 4) +
  geom_smooth(method = "lm", formula = y ~ x, color = "darkred", se = TRUE) +
  labs(
    title = "Correlation between Cloze Probability and Surprisal",
    subtitle = "Llama 3: 8B parameters",
    x = "Target Cloze Probability",
    y = "Target Surprisal (Llama 3)",
    color = "Condition"
  ) +
  # scale_color_manual(values = c("navy", "steelblue1")) +
  scale_color_manual(values = plot_colors_conditions) +
  theme_minimal() +
  theme(
    axis.title.x = element_text(size = 14, face = "bold"),
    axis.title.y = element_text(size = 14, face = "bold"),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 16, face = "italic", hjust = 0.5),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 12),
    strip.text = element_text(size = 14, face = "bold"),
    plot.background = element_rect(fill = "white", color = NA)
  )

ggsave("./plots/correlations/Surprisal_cloze_correlation_llama.png", plot = last_plot(),
       width=10, height=8, dpi=320)
```


## Correlation between Surprisal and Target RTs

```{r correlation-surprisal-rts}
dat_surprisals_mini <- dat_surprisals %>%
  filter(Region == "target") %>% 
  select("Item", "ExpCondition", "TargetWord",
         "Region", "Task", "Half",
         "Subject", "RT", "logRT",
         "ClozeProb", "TargetSurprisalGpt2Large", "TargetSurprisalGptNeo",
         "TargetSurprisalLlama")

# Correlation RTs & cloze probability (sanity check)
cor(dat_surprisals_mini$logRT,
    dat_surprisals_mini$ClozeProb)

# Correlation RTs & surprisal (sanity check)
cor(dat_surprisals_mini$logRT,
    dat_surprisals_mini$TargetSurprisalGpt2Large)

# Correlation RTs & surprisal (sanity check)
cor(dat_surprisals_mini$logRT,
    dat_surprisals_mini$TargetSurprisalGptNeo)

# Correlation RTs & surprisal (sanity check)
cor(dat_surprisals_mini$logRT,
    dat_surprisals_mini$TargetSurprisalLlama)
```

### LMER with gpt2-large surprisal as predictor

```{r surprisal-models-gpt2-large, eval = run_surprisal_models}
# Models
surprisal_models <- list()
dat_coef <- NULL
dat_surprisals$TargetSurprisalGpt2Large_c <- scale(dat_surprisals$TargetSurprisalGpt2Large)

for (region in c("target", "spill1", "spill2")) {
  
  dat_region <- dat_surprisals %>% filter(Region == region)

  m <- lmer(logRT ~ TargetSurprisalGpt2Large_c * Task +
              Half + Half:Task +
              Length_c + Freq_c +
              TrialCounterTask_c +
              logPreCritRT +
              (1 + TargetSurprisalGpt2Large_c + Task + Half | Subject) +
              (1 | Item),
            data = dat_region)
  
  print("---------------------------------------------------------------------")
  print("MODEL SUMMARY FOR REGION:")
  print(region)
  print(summary(m))
            
  coef_region <- get_coefficients(m)
  coef_region$Region <- region
  
  if (!exists("dat_coef")) {
    dat_coef <- coef_region
    } else {
    dat_coef <- rbind(dat_coef, coef_region)
  }
}

dat_coef$Region <- factor(dat_coef$Region, levels =
                            c("pre-critical","target",
                              "spill1", "spill2"))

ggplot(filter(dat_coef, Region != "pre-critical"),
       aes(x = Region, y = estimate,
           color = term)) +
  geom_errorbar(aes(ymin = estimate - std.error,
                    ymax = estimate + std.error,
                    color = term),
                width=0.1, linewidth=0.3) +
  geom_point(aes(shape = p.value < 0.05), size = 3) +
  geom_hline(yintercept = 0, color = "black") +
  scale_shape_manual(values = c(13, 19)) +
  scale_color_brewer(palette = "Set1") +
  labs(title = "Model Coefficients",
       x = "Region",
       y = "Coefficient Estimate",
       color = "Predictor") +
  theme_minimal() + 
    theme(
    axis.title.x = element_text(size = 14, face = "bold"),
    axis.title.y = element_text(size = 14, face = "bold"),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 12),
    strip.text = element_text(size = 14, face = "bold"),
    plot.background = element_rect(fill = "white", color = NA)
  )

ggsave("./plots/model_coefficients/model_coefficients_surprisal_gpt2-large.png", plot = last_plot(),
       width=10, height=7, dpi=320)
```

### LMER with GPT-Neo surprisal as predictor

```{r surprisal-models-gpt2-neo, eval = run_surprisal_models}
# Models
surprisal_models <- list()
dat_coef <- NULL
dat_surprisals$TargetSurprisalGptNeo_c <- scale(dat_surprisals$TargetSurprisalGptNeo)

for (region in c("target", "spill1", "spill2")) {
  
  dat_region <- dat_surprisals %>% filter(Region == region)

  m <- lmer(logRT ~ TargetSurprisalGptNeo_c * Task +
              Half + Half:Task +
              Length_c + Freq_c +
              TrialCounterTask_c +
              logPreCritRT +
              (1 + TargetSurprisalGptNeo_c + Task + Half | Subject) +
              (1 +  TargetSurprisalGptNeo_c + Task + Half | Item),
            data = dat_region)
  
  print("---------------------------------------------------------------------")
  print("MODEL SUMMARY FOR REGION:")
  print(region)
  print(summary(m))
            
  coef_region <- get_coefficients(m)
  coef_region$Region <- region
  
  if (!exists("dat_coef")) {
    dat_coef <- coef_region
    } else {
    dat_coef <- rbind(dat_coef, coef_region)
  }
}

dat_coef$Region <- factor(dat_coef$Region, levels =
                            c("pre-critical","target",
                              "spill1", "spill2"))

ggplot(filter(dat_coef, Region != "pre-critical"),
       aes(x = Region, y = estimate,
           color = term)) +
  geom_errorbar(aes(ymin = estimate - std.error,
                    ymax = estimate + std.error,
                    color = term),
                width=0.1, linewidth=0.3) +
  geom_point(aes(shape = p.value < 0.05), size = 3) +
  geom_hline(yintercept = 0, color = "black") +
  scale_shape_manual(values = c(13, 19)) +
  scale_color_brewer(palette = "Set1") +
  labs(title = "Model Coefficients",
       x = "Region",
       y = "Coefficient Estimate",
       color = "Predictor") +
  theme_minimal() + 
    theme(
    axis.title.x = element_text(size = 14, face = "bold"),
    axis.title.y = element_text(size = 14, face = "bold"),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 12),
    strip.text = element_text(size = 14, face = "bold"),
    plot.background = element_rect(fill = "white", color = NA)
  )

ggsave("./plots/model_coefficients/model_coefficients_surprisal_gpt-neo.png", plot = last_plot(),
       width=10, height=7, dpi=320)
```

### LMER with Llama surprisal as predictor

```{r surprisal-models-llama, eval = run_surprisal_models}
# Models
surprisal_models_llama <- list()
dat_coef <- NULL
dat_surprisals$TargetSurprisalLlama_c <- scale(dat_surprisals$TargetSurprisalLlama)

for (region in c("target", "spill1", "spill2")) {
  
  dat_region <- dat_surprisals %>% filter(Region == region)

  m <- lmer(logRT ~ TargetSurprisalLlama_c * Task +
              Half + Half:Task +
              Length_c + Freq_c +
              TrialCounterTask_c +
              logPreCritRT +
              (1 + TargetSurprisalLlama_c + Task + Half | Subject) +
              (1 | Item),
            data = dat_region)
  surprisal_models_llama <- append(surprisal_models_llama, list(m))
  
  print("---------------------------------------------------------------------")
  print("MODEL SUMMARY FOR REGION:")
  print(region)
  print(summary(m))
            
  coef_region <- get_coefficients(m)
  coef_region$Region <- region
  
  if (!exists("dat_coef")) {
    dat_coef <- coef_region
    } else {
    dat_coef <- rbind(dat_coef, coef_region)
  }
}

dat_coef$Region <- factor(dat_coef$Region, levels =
                            c("pre-critical","target",
                              "spill1", "spill2"))

ggplot(filter(dat_coef, Region != "pre-critical"),
       aes(x = Region, y = estimate,
           color = term)) +
  geom_errorbar(aes(ymin = estimate - std.error,
                    ymax = estimate + std.error,
                    color = term),
                width=0.1, linewidth=0.3) +
  geom_point(aes(shape = p.value < 0.05), size = 3) +
  geom_hline(yintercept = 0, color = "black") +
  scale_shape_manual(values = c(13, 19)) +
  scale_color_brewer(palette = "Set1") +
  labs(title = "Model Coefficients",
       x = "Region",
       y = "Coefficient Estimate",
       color = "Predictor") +
  theme_minimal() + 
    theme(
    axis.title.x = element_text(size = 14, face = "bold"),
    axis.title.y = element_text(size = 14, face = "bold"),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 12),
    strip.text = element_text(size = 14, face = "bold"),
    plot.background = element_rect(fill = "white", color = NA)
  )

ggsave("./plots/model_coefficients/model_coefficients_surprisal_llama.png", plot = last_plot(),
       width=10, height=7, dpi=320)
```


## Surprisal difference subset 

Subset for a set of items where surprisal in High < surprisal in Low
(or, if surprisal_difference > 1, additionally with a minimum distance between them):

```{r subset-strong-surprisal}
surprisal_difference <- 2.65  # resuls in 50-50 split between good vs. bad items
# surprisal_difference <- 1

dat_strong <- dat_surprisals %>% 
  dplyr::group_by(Item, ExpCondition) %>%
  dplyr::summarise(Surprisal = mean(TargetSurprisalGptNeo), .groups = "drop") %>%
  pivot_wider(names_from = ExpCondition, values_from = Surprisal) %>% # wide format
  #mutate(difference = abs(High - Low)) %>%
  dplyr::filter(High + surprisal_difference <= Low) %>%
  dplyr::select(Item) %>%
  inner_join(dat_surprisals, by = "Item")

dat_strong_surprisals <- dat_surprisals %>% 
  dplyr::filter(Item %in% dat_strong$Item) %>% 
  dplyr::arrange(Item, ExpCondition)

cat("FILTERING DATA SET FOR STRONG ITEMS ONLY\n")

cat("Subsetting data frame for", length(unique(dat_strong_surprisals$Item)),
    "unique items with a surprisal difference of at least",
    surprisal_difference, "between conditions.")
```

Print the item numbers that have lower surprisal in B than A:

```{r weak-items}
dat_weak <- dat_surprisals %>% 
  dplyr::group_by(Item, ExpCondition) %>%
  dplyr::summarise(Surprisal = mean(TargetSurprisalLlama), .groups = "drop") %>%
  pivot_wider(names_from = ExpCondition, values_from = Surprisal) %>% # wide format
  #mutate(difference = abs(High - Low)) %>%
  dplyr::filter(High >= Low) %>%
  dplyr::select(Item) %>%
  inner_join(dat_surprisals, by = "Item") %>% 
  select(Item, ExpCondition, Block) %>% 
  unique()
  
bad_items_surprisal <- droplevels(unique(dat_weak$Item))
print(bad_items_surprisal)
length(bad_items_surprisal)
```

Check for any systematic confounds of weak items with other factors:  
for example, are weak items present more in one block than another?

```{r check-weak-items}
prop.table(xtabs(~ as.character(Item) + Block, dat_weak), margin = 2)
xtabs(~Block, dat_weak) / 2
```

Yes, it looks like they are particularly present in blocks 1+3!
There are 12 problematic items in these blocks, vs. only 4 in blocks 2+4.


Optionally, look at this item set and verify that Surprisal in High is always lower than Surprisal in Low:

```{r check-strong-surprisal-item-check}
dat_strong_surprisals %>% 
  select(Item, ExpCondition, TargetWord, TargetSurprisalGptNeo, ClozeProb) %>% 
  unique()
```

Plot the RTs for this strong surprisal item subset:

```{r rt-line-plot-strong-surprisal-items, fig.height=5, fig.width=8}
dat_plot <- summarySE(dat_strong_surprisals,
                      measurevar="RT",
                      groupvars=c("Region", "ExpCondition", "Task"),
                      na.rm=T)

ggplot(data = dat_plot,
       aes(x = Region,
           y = RT,
           color = ExpCondition,
           linetype = Task,
           group = interaction(ExpCondition, Task))) +
  geom_line(linewidth = 1) +
  geom_point(size = 2, show.legend = FALSE) +
  geom_errorbar(aes(ymin = RT - se, ymax = RT + se), 
                width = 0.2, 
                linewidth = 0.5) +
  #facet_wrap(~Task) +
  xlab("Region") +
  ylab("Reading Time (ms)") +
  ggtitle("Reading Times by Task") +
  labs(color = "Expectancy") +
  # scale_color_manual(values = c("navy", "steelblue1")) +
  scale_color_manual(values = plot_colors_conditions) +
  theme_minimal() +
  theme(
    axis.title.x = element_text(size = 14, face = "bold"),
    axis.title.y = element_text(size = 14, face = "bold"),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 12),
    legend.key.width = unit(1, "cm"),
    strip.text = element_text(size = 16, face = "bold"),
    plot.background = element_rect(fill = "white", color = NA)
  )

ggsave("./plots/subset_analyses/RTs_surprisal_strong.png", plot = last_plot(),
       width=6, height=4, dpi=320)
```

Do the same but for the weak surprisal item subset (i.e., the items that were filtered out in the code above):

```{r rt-line-plot-weak-surprisal-items, fig.height=5, fig.width=8}
dat_weak_surprisals <- dat_surprisals %>% 
    filter(!(Item %in% dat_strong_surprisals$Item))

# Optionally, look at this item set
dat_weak_surprisals %>% 
  select(Item, ExpCondition, TargetWord, TargetSurprisalGptNeo, ClozeProb) %>% 
  unique()

dat_plot <- summarySE(dat_weak_surprisals,
                      measurevar="RT",
                      groupvars=c("Region", "ExpCondition", "Task"),
                      na.rm=T)

ggplot(data = dat_plot,
       aes(x = Region,
           y = RT,
           color = ExpCondition,
           linetype = Task,
           group = interaction(ExpCondition, Task))) +
  geom_line(linewidth = 1) +
  geom_point(size = 2, show.legend = FALSE) +
  geom_errorbar(aes(ymin = RT - se, ymax = RT + se), 
                width = 0.2, 
                linewidth = 0.5) +
  #facet_wrap(~Task) +
  xlab("Region") +
  ylab("Reading Time (ms)") +
  ggtitle("Reading Times by Task") +
  labs(color = "Expectancy") +
  # scale_color_manual(values = c("navy", "steelblue1")) +
  scale_color_manual(values = plot_colors_conditions) +
  theme_minimal() +
  theme(
    axis.title.x = element_text(size = 14, face = "bold"),
    axis.title.y = element_text(size = 14, face = "bold"),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 12),
    legend.key.width = unit(1, "cm"),
    strip.text = element_text(size = 16, face = "bold"),
    plot.background = element_rect(fill = "white", color = NA)
  )

ggsave("./plots/subset_analyses/RTs_surprisal_weak.png", plot = last_plot(),
       width=6, height=4, dpi=320)
```


## Low surprisal in Condition High subset

Subset the items for a low surprisal in the High expectancy condition:

```{r subset-surprisal}
max_surprisal_high <- 4.9  # resuls in 50-50 split between good vs. bad items
# max_surprisal_high <- 3

dat_surprisal_max_cond <- dat_surprisals %>% 
  filter(ExpCondition == "High" &
           TargetSurprisalGptNeo < max_surprisal_high)
dat_surprisal_max_cond <- dat_surprisals %>% 
  filter(Item %in% dat_surprisal_max_cond$Item)

length(unique(dat_surprisal_max_cond$Item))
```

Plot the RTs for this surprisal item subset with low surprisal in Condition=High:

```{r rt-line-plot-surprisal-items-high-below-threshold, fig.height=5, fig.width=8}
dat_plot <- summarySE(dat_surprisal_max_cond,
                      measurevar="RT",
                      groupvars=c("Region", "ExpCondition", "Task"),
                      na.rm=T)

ggplot(data = dat_plot,
       aes(x = Region,
           y = RT,
           color = ExpCondition,
           linetype = Task,
           group = interaction(ExpCondition, Task))) +
  geom_line(linewidth = 1) +
  geom_point(size = 2, show.legend = FALSE) +
  geom_errorbar(aes(ymin = RT - se, ymax = RT + se), 
                width = 0.2, 
                linewidth = 0.5) +
  #facet_wrap(~Task) +
  xlab("Region") +
  ylab("Reading Time (ms)") +
  ggtitle("Reading Times by Task") +
  labs(color = "Expectancy") +
  # scale_color_manual(values = c("navy", "steelblue1")) +
  scale_color_manual(values = plot_colors_conditions) +
  theme_minimal() +
  theme(
    axis.title.x = element_text(size = 14, face = "bold"),
    axis.title.y = element_text(size = 14, face = "bold"),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 12),
    legend.key.width = unit(1, "cm"),
    strip.text = element_text(size = 16, face = "bold"),
    plot.background = element_rect(fill = "white", color = NA)
  )

ggsave("./plots/subset_analyses/RTs_surprisal_high_below_threshold.png", plot = last_plot(),
       width=6, height=4, dpi=320)
```

```{r rt-line-plot-surprisal-high-above-threshold, fig.height=5, fig.width=8}
dat_surprisal_min_cond <- dat_surprisals %>% 
  filter(!(Item %in% dat_surprisal_max_cond$Item))

dat_plot <- summarySE(dat_surprisal_min_cond,
                      measurevar="RT",
                      groupvars=c("Region", "ExpCondition", "Task"),
                      na.rm=T)

ggplot(data = dat_plot,
       aes(x = Region,
           y = RT,
           color = ExpCondition,
           linetype = Task,
           group = interaction(ExpCondition, Task))) +
  geom_line(linewidth = 1) +
  geom_point(size = 2, show.legend = FALSE) +
  geom_errorbar(aes(ymin = RT - se, ymax = RT + se), 
                width = 0.2, 
                linewidth = 0.5) +
  #facet_wrap(~Task) +
  xlab("Region") +
  ylab("Reading Time (ms)") +
  ggtitle("Reading Times by Task") +
  labs(color = "Expectancy") +
  # scale_color_manual(values = c("navy", "steelblue1")) +
  scale_color_manual(values = plot_colors_conditions) +
  theme_minimal() +
  theme(
    axis.title.x = element_text(size = 14, face = "bold"),
    axis.title.y = element_text(size = 14, face = "bold"),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 12),
    legend.key.width = unit(1, "cm"),
    strip.text = element_text(size = 16, face = "bold"),
    plot.background = element_rect(fill = "white", color = NA)
  )

ggsave("./plots/subset_analyses/RTs_surprisal_high_above_threshold.png", plot = last_plot(),
       width=6, height=4, dpi=320)
```


## Full stimuli correlation analysis

Transform the data frame to extract the surprisal of every word into a new column:

```{r add-surprisal-all-words}
dat_surprisals_all$WordPosition <- as.numeric(dat_surprisals_all$WordPosition)
dat_surprisals_all <- dat_surprisals_all %>% 
  rowwise() %>% 
  mutate(Gpt2Large = as.numeric(strsplit(SurprisalsGpt2Large, " ")[[1]][WordPosition])) %>%
  mutate(GptNeo = as.numeric(strsplit(SurprisalsGptNeo, " ")[[1]][WordPosition])) %>%
  mutate(Llama = as.numeric(strsplit(SurprisalsLlama, " ")[[1]][WordPosition])) %>%
  ungroup()

# Extract surprisal at each word position
apply_surprisals_positions <- function(surprisals, pos) {
    surprisal_values <- strsplit(surprisals, " ")[[1]]  # Split into vector
    as.numeric(surprisal_values[pos])  # Extract correct index
}

dat_surprisals_all <- dat_surprisals_all %>%
  mutate(Gpt2Large = mapply(apply_surprisals_positions, SurprisalsGpt2Large, WordPosition))

dat_surprisals_all <- dat_surprisals_all %>%
  mutate(GPTNeo = mapply(apply_surprisals_positions, SurprisalsGptNeo, WordPosition))

dat_surprisals_all <- dat_surprisals_all %>%
  mutate(Llama = mapply(apply_surprisals_positions, SurprisalsLlama, WordPosition))

# delete <- dat_surprisals_all %>%
#   select(Item, ExpCondition, WordPosition, Word, Gpt2Large, SurprisalsGpt2Large) %>% 
#   arrange(Item, ExpCondition, WordPosition) %>% 
#   unique()
# 
# write.csv(delete, "delete.csv", row.names = FALSE)
```

Check the correlations between the surprisal measures themselves:

```{r cor-surprisals-all}
# Correlations among the surprisal values themselves (3 pairs)
cor(dat_surprisals_all$Gpt2Large,
    dat_surprisals_all$GPTNeo)
cor(dat_surprisals_all$Gpt2Large,
    dat_surprisals_all$Llama)
cor(dat_surprisals_all$GPTNeo,
    dat_surprisals_all$Llama)
```

As expected: largest similarity between gpt2-large and GPTNeo.


```{r cor-word-length-all}
cor(dat_surprisals_all$WordLength,
    dat_surprisals_all$Gpt2Large)
cor(dat_surprisals_all$WordLength,
    dat_surprisals_all$GPTNeo)
cor(dat_surprisals_all$WordLength,
    dat_surprisals_all$Llama)
```

Correlation keeps increasing!


Now the interesting bit: the correlation with RTs:

```{r cor-rt-all}
cor(dat_surprisals_all$RT,
    dat_surprisals_all$Gpt2Large)
cor(dat_surprisals_all$RT,
    dat_surprisals_all$GPTNeo)
cor(dat_surprisals_all$RT,
    dat_surprisals_all$Llama)

dat_surprisals_all$logRT <- log(dat_surprisals_all$RT)
cor(dat_surprisals_all$logRT,
    dat_surprisals_all$Gpt2Large)
cor(dat_surprisals_all$logRT,
    dat_surprisals_all$GPTNeo)
cor(dat_surprisals_all$logRT,
    dat_surprisals_all$Llama)
```

Why are these correlation scores so low?!


### Generate a by-item RT vs. surprisal plot.

TODO: this does not separate conditions for surprisal and does not follow the new color assignment etc. Redo the plot.

```{r item-surprisal-plots-individual, eval = FALSE}
if (plot_inidividual_items_surprisal == TRUE) {
  
  dat_plot <- summarySE(dat_surprisals_all,
                        measurevar="RT",
                        groupvars=c("Region", "ExpCondition",
                                    "Task", "Item",
                                    "WordPosition", "Word",
                                    "Gpt2Large", "GPTNeo", "Llama"),
                        na.rm=T)
  
  dat_plot$WordPosition <- as.numeric(dat_plot$WordPosition)
  summary(dat_plot$WordPosition)
  
  for (i in unique(as.numeric(as.character(dat_plot$Item)))) {
    dat_plot_item <- dat_plot %>% 
      filter(Item == i) %>% 
      group_by(WordPosition) %>%
      mutate(Word = if_else(
        WordPosition == 6,
        paste(unique(Word[WordPosition == 6]), collapse = "/"),
        Word)) %>%
      ungroup()
    item_no_char <- as.character(unique(dat_plot_item$Item))
    
    # Rescale surprisal to RTs
    dat_plot_item <- dat_plot_item %>%
      mutate(Surprisal_Gpt2Large_scaled = scales::rescale(Gpt2Large,
                                                to = range(RT, na.rm = TRUE))) %>% 
      mutate(Surprisal_GPTNeo_scaled = scales::rescale(GPTNeo,
                                             to = range(RT, na.rm = TRUE))) %>% 
      mutate(Surprisal_Llama_scaled = scales::rescale(Llama,
                                            to = range(RT, na.rm = TRUE)))
    
    dat_plot_item_long <- dat_plot_item %>%
      pivot_longer(
        cols = starts_with("Surprisal"),
        names_to = "Measure",
        values_to = "SurprisalValue"
      ) %>%
      mutate(
        Measure = recode(
          Measure,
          "Surprisal_Gpt2Large_scaled" = "GPT2-Large",
          "Surprisal_GPTNeo_scaled" = "GPTNeo",
          "Surprisal_Llama_scaled" = "Llama"
        )
      )
    
    p <- ggplot(dat_plot_item_long,
                aes(x = WordPosition)) +
      
      geom_line(aes(y = RT,
                    color = ExpCondition,
                    linetype = Task,
                    group = interaction(ExpCondition, Task)),
                linewidth = 1) +
      geom_point(aes(y = RT,
                     color = ExpCondition),
                 size = 2) +
      geom_errorbar(aes(y = RT,
                        color = ExpCondition,
                        ymin = RT - se,
                        ymax = RT + se), 
                    width = 0.2, 
                    linewidth = 0.5) +
    
      # Add scaled surprisal lines and points
      geom_line(aes(y = SurprisalValue, color = Measure),
                linewidth = 1, linetype = "solid") +
      geom_point(aes(y = SurprisalValue, color = Measure), size = 2) +

      {if (as.numeric(item_no_char) < 1000)
           annotate("rect", xmin = 5.5, xmax = 8.5, ymin = -Inf, ymax = Inf, 
               fill = "gray", alpha = 0.3)} +
      
      labs(color = "Measure",
           title = paste0("Reading Times: Item ", item_no_char)) +

      scale_x_continuous(
        name = "Word Position",
        breaks = dat_plot_item$WordPosition,
        sec.axis = sec_axis(~ ., name = "Word",
                            breaks = dat_plot_item$WordPosition,
                            labels = dat_plot_item$Word)
      ) +
      
      ylab("Reading Time (ms)") +
      # Secondary Y-Axis
      scale_y_continuous(sec.axis = sec_axis(~ ., name = "Scaled surprisal")) +

      # Custom color scale for the 3 surprisal measures
      scale_color_manual(values = c("GPT2-Large" = "gold",
                                    "GPTNeo" = "orange",
                                    "Llama" = "darkorange3",
                                    "High" = "navy",
                                    "Low" = "steelblue1")) +
      
      theme_minimal() +
      theme(
        axis.title.x = element_text(size = 14, face = "bold"),
        axis.title.y = element_text(size = 14, face = "bold"),
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
        legend.title = element_text(size = 14),
        legend.text = element_text(size = 12),
        strip.text = element_text(size = 14, face = "bold"),
        plot.background = element_rect(fill = "white", color = NA)
      ) #+
      #scale_color_manual(values = c("navy", "steelblue1"))
    
    plot_title <- paste0("./plots/item_plots_surprisal/item_", as.character(i), ".png")
    ggsave(plot_title, plot = p, width=12, height=5, dpi=320)
  }
}
```

The same but generating separate plots for production and comprehension:

TODO: This also does not yet distinguish condition or follow the newer color coding.

```{r item-surprisal-plots-individual-by-task, eval = FALSE}
if (plot_inidividual_items_surprisal == TRUE) {
  
  dat_plot <- summarySE(dat_surprisals_all,
                        measurevar="RT",
                        groupvars=c("Region", "ExpCondition",
                                    "Task", "Item",
                                    "WordPosition", "Word",
                                    "Gpt2Large", "GPTNeo", "Llama"),
                        na.rm=T)
  
  dat_plot$WordPosition <- as.numeric(dat_plot$WordPosition)
  summary(dat_plot$WordPosition)
  
  for (i in unique(as.numeric(as.character(dat_plot$Item)))) {
    
    for (task in c("Comprehension", "Production")) {
      
      dat_plot_item <- dat_plot %>% 
        filter(Item == i, Task == task) %>% 
        group_by(WordPosition) %>%
        mutate(Word = if_else(
          WordPosition == 6,
          paste(unique(Word[WordPosition == 6]), collapse = "/"),
          Word)) %>%
        ungroup()
      item_no_char <- as.character(unique(dat_plot_item$Item))
      
      # Rescale surprisal to RTs
      dat_plot_item <- dat_plot_item %>%
        mutate(Surprisal_Gpt2Large_scaled = scales::rescale(Gpt2Large,
                                                  to = range(RT, na.rm = TRUE))) %>% 
        mutate(Surprisal_GPTNeo_scaled = scales::rescale(GPTNeo,
                                               to = range(RT, na.rm = TRUE))) %>% 
        mutate(Surprisal_Llama_scaled = scales::rescale(Llama,
                                              to = range(RT, na.rm = TRUE)))
      
      dat_plot_item_long <- dat_plot_item %>%
        pivot_longer(cols = starts_with("Surprisal"),
                     names_to = "Measure",
                     values_to = "SurprisalValue") %>%
        mutate(Measure = recode(Measure,
                                "Surprisal_Gpt2Large_scaled" = "GPT2-Large",
                                "Surprisal_GPTNeo_scaled" = "GPTNeo",
                                "Surprisal_Llama_scaled" = "Llama"))
      
      p <- ggplot(dat_plot_item_long,
                  aes(x = WordPosition)) +
        
        geom_line(aes(y = RT,
                      color = ExpCondition,
                      #linetype = Task,
                      #group = interaction(ExpCondition, Task)),
                      group = ExpCondition),
                  linewidth = 1) +
        geom_point(aes(y = RT,
                       color = ExpCondition),
                   size = 2) +
        geom_errorbar(aes(y = RT,
                          color = ExpCondition,
                          ymin = RT - se,
                          ymax = RT + se), 
                      width = 0.2, 
                      linewidth = 0.5) +
      
        # Add scaled surprisal lines and points
        geom_line(aes(y = SurprisalValue, color = Measure),
                  linewidth = 1, linetype = "solid") +
        geom_point(aes(y = SurprisalValue, color = Measure), size = 2) +
  
        {if (as.numeric(item_no_char) < 1000)
             annotate("rect", xmin = 5.5, xmax = 8.5, ymin = -Inf, ymax = Inf, 
                 fill = "gray", alpha = 0.3)} +
        
        labs(color = "Measure",
             title = paste0("Reading Times: Item ", item_no_char,
                            " - : ", task)) +
  
        scale_x_continuous(
          name = "Word Position",
          breaks = dat_plot_item$WordPosition,
          sec.axis = sec_axis(~ ., name = "Word",
                              breaks = dat_plot_item$WordPosition,
                              labels = dat_plot_item$Word)
        ) +
        
        ylab("Reading Time (ms)") +
        # Secondary Y-Axis
        scale_y_continuous(sec.axis = sec_axis(~ ., name = "Scaled surprisal")) +
  
        # Custom color scale for the 3 surprisal measures
        scale_color_manual(values = c("GPT2-Large" = "gold",
                                      "GPTNeo" = "orange",
                                      "Llama" = "darkorange3",
                                      "High" = "navy",
                                      "Low" = "steelblue1")) +
        
        theme_minimal() +
        theme(
          axis.title.x = element_text(size = 14, face = "bold"),
          axis.title.y = element_text(size = 14, face = "bold"),
          axis.text.x = element_text(size = 12),
          axis.text.y = element_text(size = 12),
          plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
          legend.title = element_text(size = 14),
          legend.text = element_text(size = 12),
          strip.text = element_text(size = 14, face = "bold"),
          plot.background = element_rect(fill = "white", color = NA)
        ) #+
        #scale_color_manual(values = c("navy", "steelblue1"))
      
      plot_title <- paste0("./plots/item_plots_surprisal/item_",
                           as.character(i), "_", task, ".png")
      ggsave(plot_title, plot = p, width=12, height=5, dpi=320)
    }
  }
}
```

Try generating a separate plot for surprisal below the regular plot:

TODO: not working yet.

```{r item-surprisal-plots-individual-separate, eval = FALSE}
if (plot_inidividual_items_surprisal == TRUE) {
  
  dat_plot <- summarySE(dat_surprisals_all,
                        measurevar="RT",
                        groupvars=c("Region", "ExpCondition",
                                    "Task", "Item",
                                    "WordPosition", "Word",
                                    "Gpt2Large", "GPTNeo", "Llama"),
                        na.rm=T)
  
  dat_plot$WordPosition <- as.numeric(dat_plot$WordPosition)
  summary(dat_plot$WordPosition)
  
  for (i in unique(as.numeric(as.character(dat_plot$Item)))) {
    dat_plot_item <- dat_plot %>% 
      filter(Item == i) %>% 
      group_by(WordPosition) %>%
      mutate(Word = if_else(
        WordPosition == 6,
        paste(unique(Word[WordPosition == 6]), collapse = "/"),
        Word)) %>%
      ungroup() %>% 
      dplyr::rename(Measure_RT = RT,
                    Measure_Gpt2Large = Gpt2Large,
                    Measure_GPTNeo = GPTNeo,
                    Measure_Llama = Llama)
    item_no_char <- as.character(unique(dat_plot_item$Item))
    
    # Rescale surprisal to RTs
    # dat_plot_item <- dat_plot_item %>%
    #   mutate(Surprisal_Gpt2Large_scaled = scales::rescale(Gpt2Large,
    #                                             to = range(RT, na.rm = TRUE))) %>% 
    #   mutate(Surprisal_GPTNeo_scaled = scales::rescale(GPTNeo,
    #                                          to = range(RT, na.rm = TRUE))) %>% 
    #   mutate(Surprisal_Llama_scaled = scales::rescale(Llama,
    #                                         to = range(RT, na.rm = TRUE)))
    
    dat_plot_item_long <- dat_plot_item %>%
      pivot_longer(cols = starts_with("Measure"),
                   names_to = "Measure",
                   values_to = "Value") %>%
      mutate(Measure = recode(Measure,
                          "Measure_RT" = "RT",
                          "Measure_Gpt2Large" = "Gpt2Large",
                          "Measure_GPTNeo" = "GPTNeo",
                          "Measure_Llama" = "Llama"))
    dat_plot_item_long$MeasureType = ifelse(dat_plot_item_long$Measure == "RT",
                                            "RT", "Surprisal")
    
    p <- ggplot(dat_plot_item_long,
                aes(x = WordPosition)) +
      
      geom_line(aes(y = Value,
                    color = ExpCondition,
                    linetype = Task,
                    group = interaction(ExpCondition, Task)),
                linewidth = 1) +
      geom_point(aes(y = Value,
                     color = ExpCondition),
                 size = 2) +
      geom_errorbar(aes(y = Value,
                        color = ExpCondition,
                        ymin = Value - se,
                        ymax = Value + se), 
                    width = 0.2, 
                    linewidth = 0.5) +
      
      facet_wrap(~MeasureType, nrow=2) +
    
      # # Add scaled surprisal lines and points
      # geom_line(aes(y = SurprisalValue, color = Measure),
      #           linewidth = 1, linetype = "solid") +
      # geom_point(aes(y = SurprisalValue, color = Measure), size = 2) +

      {if (as.numeric(item_no_char) < 1000)
           annotate("rect", xmin = 5.5, xmax = 8.5, ymin = -Inf, ymax = Inf, 
               fill = "gray", alpha = 0.3)} +
      
      labs(color = "Measure",
           title = paste0("Reading Times: Item ", item_no_char)) +

      scale_x_continuous(
        name = "Word Position",
        breaks = dat_plot_item$WordPosition,
        sec.axis = sec_axis(~ ., name = "Word",
                            breaks = dat_plot_item$WordPosition,
                            labels = dat_plot_item$Word)
      ) +
      
      ylab("Reading Time (ms)") +
      # Secondary Y-Axis
      #scale_y_continuous(sec.axis = sec_axis(~ ., name = "Scaled surprisal")) +

      # Custom color scale for the 3 surprisal measures
      # scale_color_manual(values = c("GPT2-Large" = "gold",
      #                               "GPTNeo" = "orange",
      #                               "Llama" = "darkorange3",
      #                               "High" = "navy",
      #                               "Low" = "steelblue1")) +
      
      scale_color_manual(values = c("navy", "steelblue1")) +
      theme_minimal() +
      theme(
        axis.title.x = element_text(size = 14, face = "bold"),
        axis.title.y = element_text(size = 14, face = "bold"),
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
        legend.title = element_text(size = 14),
        legend.text = element_text(size = 12),
        strip.text = element_text(size = 14, face = "bold"),
        plot.background = element_rect(fill = "white", color = NA)
      )
    
    plot_title <- paste0("./plots/item_plots_surprisal/item_", as.character(i), ".png")
    ggsave(plot_title, plot = p, width=12, height=5, dpi=320)
  }
}
```

Generate a separate surprisal plot for each item:

```{r item-surprisal-plots-indiviual-surprisalonly}
# if (plot_inidividual_items_surprisal == TRUE) {
#   
#   dat_plot <- summarySE(dat_surprisals_all,
#                         measurevar="RT",
#                         groupvars=c("Region", "ExpCondition",
#                                     "Task", "Item",
#                                     "WordPosition", "Word",
#                                     "Gpt2Large", "GPTNeo", "Llama"),
#                         na.rm=T)
#   
#   dat_plot$WordPosition <- as.numeric(dat_plot$WordPosition)
#   summary(dat_plot$WordPosition)
#   
#   for (i in unique(as.numeric(as.character(dat_plot$Item)))) {
#     dat_plot_item <- dat_plot %>% 
#       filter(Item == i) %>% 
#       group_by(WordPosition) %>%
#       mutate(Word = if_else(
#         WordPosition == 6,
#         paste(unique(Word[WordPosition == 6]), collapse = "/"),
#         Word)) %>%
#       ungroup() %>% 
#       dplyr::rename(Surprisal_Gpt2Large = Gpt2Large,
#                     Surprisal_GPTNeo = GPTNeo,
#                     Surprisal_Llama = Llama)
#     item_no_char <- as.character(unique(dat_plot_item$Item))
#     
#     dat_plot_item_long <- dat_plot_item %>%
#       pivot_longer(cols = starts_with("Surprisal"),
#                    names_to = "Model",
#                    values_to = "Surprisal") %>%
#       mutate(Model = recode(Model,
#                             "Surprisal_Gpt2Large" = "Gpt2Large",
#                             "Surprisal_GPTNeo" = "GPTNeo",
#                             "Surprisal_Llama" = "Llama"))
#     
#     p <- ggplot(dat_plot_item_long,
#                 aes(x = WordPosition)) +
#       
#       geom_line(aes(y = Surprisal,
#                     #color = interaction(ExpCondition, Model),
#                     color = Model,
#                     linetype = ExpCondition,
#                     group = interaction(ExpCondition, Model)),
#                 linewidth = 1) +
#       # geom_point(aes(y = Surprisal), size = 2) +
# 
#       {if (as.numeric(item_no_char) < 1000)
#            annotate("rect", xmin = 5.5, xmax = 8.5, ymin = -Inf, ymax = Inf, 
#                fill = "gray", alpha = 0.3)} +
#       
#       labs(color = "Model",
#            linetype = "Expectancy",
#            title = paste0("Surprisal values: Item ", item_no_char)) +
# 
#       scale_x_continuous(
#         name = "Word Position",
#         breaks = dat_plot_item$WordPosition,
#         sec.axis = sec_axis(~ ., name = "Word",
#                             breaks = dat_plot_item$WordPosition,
#                             labels = dat_plot_item$Word)
#       ) +
#       
#       ylab("Surprisal") +
# 
#       # Custom color scale for the 3 surprisal measures & conditions
#       # scale_color_manual(values = c("High.Gpt2Large" = "gold3",
#       #                               "Low.Gpt2Large" = "gold1",
#       #                               "High.GPTNeo" = "orange3",
#       #                               "Low.GPTNeo" = "orange",
#       #                               "High.Llama" = "firebrick3",
#       #                               "Low.Llama" = "firebrick1")) +
#       # scale_color_manual(values = c("High.Gpt2Large" = "lightseagreen",
#       #                               "Low.Gpt2Large" = "powderblue",
#       #                               "High.GPTNeo" = "darkorchid4",
#       #                               "Low.GPTNeo" = "darkorchid3",
#       #                               "High.Llama" = "red3",
#       #                               "Low.Llama" = "red1")) +
#       scale_color_manual(values = c("Gpt2Large" = "gold",
#                                     "GPTNeo" = "orange2",
#                                     "Llama" = "firebrick")) +
# 
#       theme_minimal() +
#       theme(
#         axis.title.x = element_text(size = 14, face = "bold"),
#         axis.title.y = element_text(size = 14, face = "bold"),
#         axis.text.x = element_text(size = 12),
#         axis.text.y = element_text(size = 12),
#         plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
#         legend.title = element_text(size = 14),
#         legend.text = element_text(size = 12),
#         strip.text = element_text(size = 14, face = "bold"),
#         plot.background = element_rect(fill = "white", color = NA)
#       )
#     
#     plot_title <- paste0("./plots/item_plots_surprisal/item_",
#                          as.character(i), "_surprisals", ".png")
#     ggsave(plot_title, plot = p, width=12, height=5, dpi=320)
#   }
# }


if (plot_inidividual_items_surprisal == TRUE) {
  
  dat_plot <- summarySE(dat_surprisals_all,
                        measurevar="RT",
                        groupvars=c("Region", "ExpCondition",
                                    "Task", "Item",
                                    "WordPosition", "Word",
                                    "Gpt2Large", "GPTNeo", "Llama"),
                        na.rm=T)
  
  dat_plot$WordPosition <- as.numeric(dat_plot$WordPosition)
  summary(dat_plot$WordPosition)
  
  for (i in unique(as.numeric(as.character(dat_plot$Item)))) {
    dat_plot_item <- dat_plot %>% 
      filter(Item == i) %>% 
      group_by(WordPosition) %>%
      mutate(Word = if_else(
        WordPosition == 6,
        paste(unique(Word[WordPosition == 6]), collapse = "/"),
        Word)) %>%
      ungroup() %>% 
      dplyr::rename(Surprisal_Gpt2Large = Gpt2Large,
                    Surprisal_GPTNeo = GPTNeo,
                    Surprisal_Llama = Llama)
    item_no_char <- as.character(unique(dat_plot_item$Item))
    
    dat_plot_item_long <- dat_plot_item %>%
      pivot_longer(cols = starts_with("Surprisal"),
                   names_to = "Model",
                   values_to = "Surprisal") %>%
      mutate(Model = recode(Model,
                            "Surprisal_Gpt2Large" = "Gpt2Large",
                            "Surprisal_GPTNeo" = "GPTNeo",
                            "Surprisal_Llama" = "Llama"))
    
    p <- ggplot(dat_plot_item_long,
                # aes(x = WordPosition)) +
                # Minimally offset along the x-axis to avoid that the 
                # conditions completely overlap until the target:
                aes(x = WordPosition + (as.numeric(factor(ExpCondition)) - 1.5) * 0.04)) +
      
      geom_line(aes(y = Surprisal,
                    color = ExpCondition,
                    linetype = Model,
                    group = interaction(ExpCondition, Model)),
                linewidth = 1) +
      geom_point(aes(y = Surprisal,
                     color = ExpCondition), size = 2) +

      {if (as.numeric(item_no_char) < 1000)
           annotate("rect", xmin = 5.5, xmax = 8.5, ymin = -Inf, ymax = Inf, 
               fill = "gray", alpha = 0.3)} +
      
      labs(color = "Expectancy",
           linetype = "Model",
           title = paste0("Surprisal values: Item ", item_no_char)) +

      scale_x_continuous(
        name = "Word Position",
        breaks = dat_plot_item$WordPosition,
        sec.axis = sec_axis(~ ., name = "Word",
                            breaks = dat_plot_item$WordPosition,
                            labels = dat_plot_item$Word)
      ) +
      
      ylab("Surprisal") +

      # scale_color_manual(values = c("navy", "steelblue1")) +
      scale_color_manual(values = plot_colors_conditions) +
      
      scale_linetype_manual(values = c("Gpt2Large" = "dotted",
                                       "GPTNeo" = "dashed",
                                       "Llama" = "solid")) +

      theme_minimal() +
      theme(
        axis.title.x = element_text(size = 14, face = "bold"),
        axis.title.y = element_text(size = 14, face = "bold"),
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
        legend.title = element_text(size = 14),
        legend.text = element_text(size = 12),
        strip.text = element_text(size = 14, face = "bold"),
        plot.background = element_rect(fill = "white", color = NA)
      )
    
    plot_title <- paste0("./plots/item_plots_surprisal/item_",
                         as.character(i), "_surprisals", ".png")
    ggsave(plot_title, plot = p, width=12, height=5, dpi=320)
  }
}
```


