---
title: "Data preprocessing"
author: "Miriam Schulz"
date: "2024-11-20"
output:
  html_document:
    number_sections: true
    toc: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# About

This script preprocesses the raw PC Ibex results of the Cloze prediction & production experiment.

What this script does:

- summary of demographics
- summary of post-experimental survey
- preprocess the reading times (filter columns, assign correct column names, etc.)
- remove participants that must be excluded (due to failure to understand the task or not being native English speakers)
- check the remaining data for completeness, accuracy on the comprehension questions, number of timeouts per participant during the production task, and mean question answer time per condition
- remove outliers:
  - by a hard threshold
  - with a by-subject standard deviation approach
  - optional: removing subjects with very fast RTs in at least one task
  - optional: removing subjects with invariant RTs between tasks / with invariant or higher RTs in comprehension
- generate a bar plot of the remaining RTs by region
- export the remaining data as two R data sets (one with all regions and one with only the target regions)


# Preliminaries

Libraries & co:

```{r preliminaries, message = FALSE}
rm(list = ls())
library(tidyverse)
library(ggplot2)
library(Rmisc)
source("functions.R")
```

## Set flags

```{r flags}
exclude_participants = TRUE
exclude_fast_participants = FALSE
exclude_invariant_participants = FALSE
generate_main_plots = TRUE
```

## Read data

```{r read-data}
dat <- read.pcibex('results.csv')
dat <- unique(dat)
```


# Demographics (all participants)

## General demographics

```{r demopgrahics}
dat.demogs <- dat %>%
  filter(PennElementName == 'demographics' | Parameter == "prolific_id") %>% 
  select(UniqueID, Parameter, Value)

# Fix whitespace characters
dat.demogs$Value <- gsub('%2C', ',', dat.demogs$Value)   # commas instead of %2C
dat.demogs$Value <- gsub('%0A', ' ', dat.demogs$Value)   # spaces instead of %0A

# Transform to wide format
dat.demogs <- dat.demogs %>%
  pivot_wider(names_from = Parameter, values_from = Value)

# Make all variables factors and then age numeric
dat.demogs[] <- lapply(dat.demogs, tolower)
dat.demogs[-3] <- lapply(dat.demogs[-3], as.factor)  # all columns except Age
dat.demogs$Age <- as.numeric(dat.demogs$Age)

# Native languages
summary(dat.demogs$Native_language)

# Age
mean(dat.demogs$Age)
range(dat.demogs$Age)

# Gender
summary(dat.demogs$Gender)
```


## Prolific completion times

```{r completion-time}
# Add the prolific IDs
demogs.prolific <- read.csv("demographics_prolific.csv", header=TRUE)
demogs.prolific <- demogs.prolific %>%
  dplyr::rename(prolific_id = Participant.id,
                Birthplace_prolific = Country.of.birth,
                Time_taken = Time.taken) %>% 
  dplyr::filter(Status != "RETURNED") %>% 
  dplyr::select(prolific_id, Birthplace_prolific, Time_taken)

dat.demogs <- merge(dat.demogs, demogs.prolific, by="prolific_id", all.x=TRUE)
dat.demogs$Time_taken_h <- convert_to_hours(dat.demogs$Time_taken)

# Calculate how the non-natives influenced the median of time taken
convert_to_hours(median(dat.demogs$Time_taken))
non_natives <- c("vuqqnd", "ee49u2", "vznkz3", "3jw4ew", "ps4dv4")

# Median completion time of the non-natives:
dat.demogs %>%
  filter(UniqueID %in% non_natives) %>% 
  summarise(median_time = median(Time_taken, na.rm = TRUE),
            mean_time = mean(Time_taken, na.rm = TRUE)) %>%
  mutate(median_time_hours = convert_to_hours(median_time),
         mean_time_hours = convert_to_hours(mean_time))

# Median completion time of the natives:
dat.demogs %>%
  filter(!UniqueID %in% non_natives) %>% 
  summarise(median_time = median(Time_taken, na.rm = TRUE),
            mean_time = mean(Time_taken, na.rm = TRUE)) %>%
  mutate(median_time_hours = convert_to_hours(median_time),
         mean_time_hours = convert_to_hours(round(mean_time)))
```

```{r export-demographics}
write.csv(dat.demogs, "demographics.csv", row.names=FALSE)
```


# Post-experimental survey

```{r post-experimental-survey}
dat.survey <- dat %>%
  filter(PennElementName == 'postexp_survey') %>% 
  select(UniqueID, Parameter, Value)
dat.survey$Value <- gsub('%2C', ',', dat.survey$Value)   # commas instead of %2C
dat.survey$Value <- gsub('%0A', ' ', dat.survey$Value)   # spaces instead of %0A

# Transform to wide format
dat.survey <- dat.survey %>%
  pivot_wider(names_from = Parameter, values_from = Value)

dat.survey[] <- lapply(dat.survey, as.factor)
summary(dat.survey$Experiment_length)
summary(dat.survey$Task_difficulty)

write.csv(dat.survey, "postexp_survey.csv", row.names=FALSE)
```


# Preprocess RT data frame

```{r preprocess-rt-data}
# Fix comma character
dat$Value <- gsub('%2C', ',', dat$Value)   # commas instead of %2C
dat$Comments <- gsub('%2C', ',', dat$Comments)   # commas instead of %2C

# Keep only task data
dat <- dat %>% 
  filter(Task %in% c("Comprehension", "Production"))

# Rename columns
new_colnames <- c("WordPosition", "Word", "EventTime", "UniqueID",
                  "RandomOrder", "LatinList", "LatinListBinary",
                  "TaskOrder", "BlocksReversed",
                  "Block", "Task",
                  "ExpItemNum", "ExpItemNumOriginal",
                  "ExpItemType", "ExpCondition",
                  "TargetPosition", "TargetWord", "ClozeProb",
                  "TargetFreq", "TargetLength", "ContextNoun",
                  "SentenceEnd",
                  "TrialCounterGlobal", "TrialCounterTask",
                  "AnswerTime",
                  "TargetAnswer", "QuestionText", "CorrectAnswer",
                  "RunningAccuracy", "ProductionTimeout", "RecordingFilename",
                  "RT", "Sentence")
dat$Newline. <- NULL
dat <- select(dat, 10:42)
colnames(dat) <- new_colnames

dat$Sentence <- gsub('%2C', ',', dat$Sentence)   # commas instead of %2C

# Reorder columns
new_colnames <- c("EventTime", "UniqueID",
                  "RandomOrder", "LatinList", "LatinListBinary",
                  "TrialCounterGlobal", "TrialCounterTask",
                  "TaskOrder", "BlocksReversed",
                  "Block", "Task",
                  "ExpItemNum", "ExpItemNumOriginal",
                  "ExpItemType", "ExpCondition",
                  "WordPosition", "Word", "TargetWord", "ClozeProb",
                  "RT", "AnswerTime",
                  "TargetPosition", 
                  "TargetFreq", "TargetLength", "ContextNoun",
                  "TargetAnswer", "CorrectAnswer", "RunningAccuracy",
                  "QuestionText",
                  "ProductionTimeout",
                  "RecordingFilename",
                  "Sentence", "SentenceEnd")
dat <- dat[, new_colnames]

# Export
# write.csv(dat, "results_preprocessed.csv", row.names=FALSE)
```

Annotate the critical regions etc:

```{r preprocess-data}
dat <- preprocess_data(dat)
```


# Remove participants

Exclude subjects who were not native speakers of English or failed the tasks:

```{r exclude-participants}
if (exclude_participants == TRUE) {
  dat <- dat %>%
    # nonnatives Prolific Exp 1
    filter(!UniqueID %in% c(# FIRST PROLIFIC EXPERIMENT:
                            "vuqqnd",  # non-native, super slow RTs
                            "ee49u2",  # non-native
                            "vznkz3",  # non-native?, some double submissions (technical problems)
                            "3jw4ew",  # non-native
                            "ps4dv4",  # non-native, did not understand the task
                            # SECOND PROLIFIC EXPERIMENT:
                            "2jcruh",  # non-native and makes a funny voice on many trials
                            "e2zqww",  # read contexts aloud 
                            "0jxphp",  # empty recordings, ~50% acc, super fast RTs [RETURNED]
                            "m7ugwz",  # read contexts aloud (but faster than in comprehension)
                            "jocpf0",  # non-native and timed out (restarted)
                            "pizxpx",  # empty recordings
                            "hpttkp",  # nonsense and repeating answers [RETURNED]
                            "icak56",  # non-native??
                            "8v42p8",  # non-native?, very slow on production RTs (~1000ms)
                            "bdhmxy",  # non-native, very slow RTs in both tasks
                            "glbu6a",  # non-native, repeated last context words, wrong browser [RETURNED]
                            "6wqbuo",  # non-native
                            "-u775t",  # technical problems: restarted 3x; loud background talking
                            "ev0tqj",  # read contexts aloud in one prod block 
                            "4ftnhn",  # no server data; born in Ghana; completion time too long
                            "teg94q")  # non-native, comp accuracy ~50%, too many timeouts
    )
}
```


## TODO: new demopgrahics of keep subjects


# Checks

## Data completeness check

```{r completeness-checks}
# Subset the data frame for checks
completeness <- dat %>%
  select(UniqueID, Item, ExpCondition, Task, Block, RandomOrder, LatinList) %>%
  unique()

# All items: How often was each item presented? 
range(xtabs(~ Item, completeness))  # should show N participants

# Items by subject: did every subject see every item?
range(xtabs(~ UniqueID + Item, completeness))  # should range from 1 to 1

# Experimental conditions per subject (should be 40 for each condition)
xtabs(~ UniqueID + ExpCondition, droplevels(filter(dat, Region == "target" &
                                            ExpCondition %in% c("High", "Low"))))

# Task by subject
xtabs(~ UniqueID + Task, completeness)  # should show 80 for each task

# Block by subject
xtabs(~ UniqueID + Block, completeness)  # should show 40 for each block

# List by subject
# xtabs(~ UniqueID + LatinList, completeness)  # must show 160 or 0 everywhere
completeness %>%
  select(LatinList, UniqueID) %>%
  arrange(LatinList, UniqueID) %>% 
  unique()

# Random order
# should show 1 or 0; values > 1 mean some subjects ran the same order
xtabs(~ RandomOrder + LatinList, completeness) / 160
range(xtabs(~ RandomOrder + LatinList, completeness) / 160)

# Random order sorted by subject
completeness %>%
  select(UniqueID, RandomOrder) %>%
  arrange(UniqueID) %>% 
  unique()

# Random Order by sorted by List
completeness %>% 
  select(LatinList, RandomOrder, UniqueID) %>% 
  unique() %>% 
  arrange(LatinList, RandomOrder)

# Random order by subject+list (useful to get the latest RandomOrder used, i.e. counter state):
completeness %>% 
  select(LatinList, RandomOrder, UniqueID) %>% 
  unique()

# Item by list
range(xtabs(~ Item + LatinList, completeness)) # N subjects per list: min to max

# Subjects per list
n_trials <- 160
xtabs(~ LatinList, completeness) / n_trials
```


## Comprehension question accuracy

```{r question-accuracy}
dat_acc <- dat %>%
  filter(!is.na(CorrectAnswer)) %>%
  select(CorrectAnswer, UniqueID, LatinList, Item) %>% 
  unique()
  #filter(Region == "target")
dat_acc$TrialAccuracy <- ifelse(dat_acc$CorrectAnswer == "true", 1, 0)

# Completeness checks
table(xtabs(~ UniqueID, dat_acc)) # should equal N of comp trials with Q in the list & N subjects

# Acc for all trials
dat_acc_subj <- dat_acc %>%
  dplyr::group_by(UniqueID, LatinList) %>%
  dplyr::summarize(Accuracy = mean(TrialAccuracy))
# dat_acc_subj
round(range(dat_acc_subj$Accuracy), 2)
round(mean(dat_acc_subj$Accuracy), 2)
```

As a sanity check: The calculated accuracy should equal the accuracy as calculated live during the experiment up to the last comprehension trial.

```{r}
dat %>%
  filter(Task == "Comprehension") %>% 
  filter(TrialCounterTask == max(TrialCounterTask)) %>% 
  select(UniqueID, RunningAccuracy) %>% 
  unique() %>% 
  mutate(RunningAccuracy = round(RunningAccuracy, 3)) %>%
  arrange(UniqueID)
```

Plot the accuracy for fast visual inspection:

```{r accuracy-subjects-plot, eval = generate_main_plots, fig.height=7, fig.width=10}
dat_acc_subj$UniqueID <- factor(dat_acc_subj$UniqueID, 
                                levels = dat_acc_subj$UniqueID[order(dat_acc_subj$Accuracy,
                                                                     dat_acc_subj$LatinList,
                                                                     decreasing = TRUE)])

ggplot(data=dat_acc_subj,
       aes(UniqueID, Accuracy, fill=as.factor(LatinList))) + 
  geom_bar(stat="identity",
           width=1,
           show.legend=TRUE,
           color = "gray",
           alpha = 0.5) +
  geom_hline(yintercept = 0.7, linetype = "dashed", color = "black",
             linewidth = 0.8) +
  annotate("text",
           x = length(unique(dat_acc_subj$UniqueID)) / 2, y = 0.75,
           label = "70% accuracy",
           color = "black", size = 8, fontface = "bold",
           angle = 0) +
  labs(title = "Mean accuracy by subject", 
       subtitle = "All experimental trials, practice excluded", 
       x = "Subject", 
       y = "Mean accuracy",
       fill = "List") +
  scale_fill_manual(values=c("plum4", "steelblue4", "gray40", "darkred",
                             "plum2", "lightsteelblue", "gray70",
                             "lightcoral")) +
  theme_minimal() +
  theme(
    axis.title.x = element_text(size = 14, face = 'bold'),
    axis.title.y = element_text(size = 14, face = 'bold'),
    axis.text.x = element_text(size = 12, face = 'bold', angle = 90,
                               vjust = 0.5, hjust = 1),
    axis.text.y = element_text(size = 12),
    plot.title = element_text(size = 25, face = 'bold', hjust = 0.5),
    plot.subtitle = element_text(size = 16, face = "italic", hjust = 0.5)
  )
```


## Production timeouts

Plot the number of production timeouts by participant:

```{r timeout-subjects-plot, eval = generate_main_plots, fig.height=5, fig.width=15}
dat_Timeouts <- dat %>% 
  filter(Task == "Production") %>% 
  select(c("UniqueID", "ExpCondition", "ProductionTimeout",
           "ExpItemNum", "AnswerTime")) %>% 
  unique()
dat_Timeouts$ProductionTimeoutBinary <-
  ifelse(dat_Timeouts$ProductionTimeout == "TimedOut", 1, 0)

ggplot(data=dat_Timeouts,
       aes(UniqueID, ProductionTimeoutBinary, fill=UniqueID)) + 
  geom_bar(stat="identity",
           width=1,
           show.legend=FALSE) +
  xlab("Subject") +
  ylab("N Production Timeouts") +
  ggtitle("Production Timeouts by Subject") +
  theme_minimal() +
  theme(
    axis.title.x = element_text(size = 14, face = 'bold'),
    axis.title.y = element_text(size = 14, face = 'bold'),
    axis.text.x = element_text(size = 12, face = 'bold', angle = 90,
                               vjust = 0.5, hjust = 1, margin = margin(t = 5)),
    axis.text.y = element_text(size = 12),
    plot.title = element_text(size = 25, face = 'bold', hjust = 0.5),
    plot.subtitle = element_text(size = 18, face = "italic", hjust = 0.5)
  )
```


## Question answer time

Comprehension question Answer Time:

```{r question-answertime-plots, eval = generate_main_plots}
dat_AnswerRTs <- dat %>%
  filter(QuestionText != "")

# Plot question answer times by subject
ggplot(data=dat_AnswerRTs,
        aes(AnswerTime, fill=UniqueID)) + 
  geom_histogram(alpha = 0.5,
                 show.legend=TRUE) + 
  labs(title = "Question RTs by Subject", 
       subtitle = "All conditions (including fillers)", 
       fill = "Subject") +
  theme_minimal() +
  theme(
    axis.title.x = element_text(size = 14, face = 'bold'),
    axis.title.y = element_text(size = 14, face = 'bold'),
    axis.text.x = element_text(size = 12, face = 'bold'),
    axis.text.y = element_text(size = 12),
    plot.title = element_text(size = 25, face = 'bold', hjust = 0.5),
    plot.subtitle = element_text(size = 16, face = "italic", hjust = 0.5),
    strip.text = element_text(size = 18, face = "bold")
  ) 

# Filter out Answer Times shorter than 200ms and longer than 10s (arbitrary):
dat_AnswerRTs <- dat_AnswerRTs %>%
  filter(AnswerTime > 200 & AnswerTime < 10000)

range(dat_AnswerRTs$AnswerTime)
mean(dat_AnswerRTs$AnswerTime)
sd(dat_AnswerRTs$AnswerTime)

# Plot question answer times by condition
ggplot(data=filter(dat_AnswerRTs, ExpCondition %in% c("High", "Low")),
        aes(AnswerTime, fill=ExpCondition)) + 
  geom_density(alpha = 0.7,
                 show.legend=TRUE) + 
  labs(title = "Question RTs by Condition", 
       subtitle = "Experimental conditions only; RTs filtered > 200ms & < 10s", 
       fill = "Subject") +
  scale_fill_manual(values=c("navy", "steelblue1")) +
  theme_minimal() +
  theme(
    axis.title.x = element_text(size = 14, face = 'bold'),
    axis.title.y = element_text(size = 14, face = 'bold'),
    axis.text.x = element_text(size = 12, face = 'bold'),
    axis.text.y = element_text(size = 12),
    plot.title = element_text(size = 25, face = 'bold', hjust = 0.5),
    plot.subtitle = element_text(size = 16, face = "italic", hjust = 0.5),
    strip.text = element_text(size = 18, face = "bold")
  ) 
```


# Outlier elimination

To continue, keep only the precritical regions and regions of interest, and split the data into all and experimental items only:

```{r filter-regions-fillers}
dat_all <- dat %>% 
  filter(!is.na(RT)) %>% 
  filter(Word != "")
dat <- dat %>% 
  filter(Region != "other")
dat <- dat %>%
  filter(dat$ExpItemType == "ExpItem")
```

Visually check for outliers:

```{r plot-rts-before-outlier-elimination, eval = generate_main_plots}
rt_hist(dat, plot_subtitle = "All critical regions, before outlier elimination")
```

## Remove super fast subjects (optional)

First, optionally exclude subjects whose mean RT was < 200ms (or some other threshold) in at least one of the two tasks:

```{r remove-fast-subjects}
if (exclude_fast_participants == TRUE) {
  dat <- remove_fast_subjects(dat, threshold = 200)
  dat_all <- remove_fast_subjects(dat_all, threshold = 200)
}
```

## Threshold based outlier elimination

Remove trials with extreme RTs:

```{r remove-outliers-threshold}
dat <- remove_outliers_thresholds(dat, min_rt = 100, max_rt = 2500)
dat_all <- remove_outliers_thresholds(dat_all, min_rt = 100, max_rt = 2500)
```

## SD by subject based outlier elimination

Additional outlier elimination by SD by participant:

```{r remove-outliers-sd}
dat <- remove_outliers_subj_sd(dat, sd_threshold=2.5)
dat_all <- remove_outliers_subj_sd(dat_all, sd_threshold=2.5)
```

Check region times after outlier elimination:

```{r print-rts-after-outlier-elimination}
print_RTs(dat)
```

```{r plot-rts-after-outlier-elimination, eval = generate_main_plots}
rt_hist(dat, plot_subtitle = "All critical regions, after outlier elimination")
```


## Remove invariant participants (optional)

Get participants who were invariant between the production / comprehension task (i.e., for which the difference between production wrt. comprehension is <= 0, or <= 10 on average):

```{r remove-invariant-subjects}
if (exclude_invariant_participants == TRUE) {
  dat <- remove_invavriant_subjects(dat, leniency = 10,
                                    remove_which = "invariant")
  dat_all <- remove_invavriant_subjects(dat_all, leniency = 10,
                                        remove_which = "invariant")
}
```



# Bar plot of RTs

```{r rt-bar-plot, eval = generate_main_plots, fig.height=5, fig.width=12}
# Generate a summary data frame including the SE:
dat_plot <- summarySE(filter(dat, !(Region %in% c("other"))),
                      measurevar="RT",
                      groupvars=c("ExpCondition", "Task", "Region"),
                      na.rm=T)

# Determine maximal value for y-axis:
max_y <- max(c(dat_plot$RT + dat_plot$se))
dat_plot$Region <- as.factor(dat_plot$Region)

# Plot:
ggplot(data=dat_plot,
       # aes(ExpCondition, RT, fill=ExpCondition)) +
       aes(ExpCondition, RT, fill = interaction(Task, ExpCondition))) + 
  geom_bar(#aes(fill=ExpCondition,
               # alpha = Task == "Production"
               # ),
           stat="identity",
           width=1,
           show.legend=TRUE) +
  geom_errorbar(aes(ymin = RT-se, ymax = RT+se),
              width=0.1, linewidth=0.3) +
  facet_wrap(~Region+Task, nrow = 1) +
  xlab("Condition") +
  ylab("Reading Time (ms)") +
  ggtitle("RTs by Task and Region") +
  # scale_fill_manual(name = "Condition",
  #                   values=c("navy", "steelblue1")) +
  scale_fill_manual(
    name = "Task and Condition", # Legend title
    values = c("Comprehension.High" = "navy", 
               "Comprehension.Low" = "lightsteelblue1",
               "Production.High" = "seagreen4",
               "Production.Low" = "#d5f5df"),
    labels = c("Comprehension: High", "Production: High",
               "Comprehension: Low", "Production: Low")
  ) +
  ylim(0, max_y) +
  theme_minimal() + 
  theme(
      axis.title.x = element_text(size = 14, face = 'bold'),
      axis.title.y = element_text(size = 14, face = 'bold'),
      axis.text.x = element_text(size = 12, face = 'bold'),
      axis.text.y = element_text(size = 12),
      plot.title = element_text(size = 25, face = 'bold', hjust = 0.5),
      #legend.position = "none",
      strip.text = element_text(size = 18, face = "bold"),
      plot.background = element_rect(fill = "white", color = NA)
  )

ggsave("./plots/RTs_barplot.png", plot = last_plot(),
       width=18, height=6, dpi=320)
```

Numeric check of RTs:

```{r check-rts}
print_RTs(dat)
```


# Export

```{r export-r-datasets}
# write.csv(dat, "results_preprocessed_filtered_targetregions.csv", row.names=FALSE)
# write.csv(dat_all, "results_preprocessed_filtered_all.csv", row.names=FALSE)
save(dat, file = "results_preprocessed_targetregions.RData")
save(dat_all, file = "results_preprocessed_allregions.RData")
```

